Today was not going well for Linda Greene.
It had started out alright.
A brisk walk in the frosty air at sunrise, a quick trip up to the castle to drop off a couple of poultices for the servants there.
The cook had even given her a big side of braised ham for her help.
But things had gone downhill pretty quickly when the warty old crone had strolled in and started turning people into frogs.

Now here she was, speeding over the mountaintops, hair and cloak whipped back by the frozen wind, and a crazy old hag hot on her tail.
The crone had a wicked-looking knife clutched between her teeth.
Been screaming that she was going to gut the king with it, or something unpleasant like that.
Well, the king was safe for now, even if he was croaking rather indignantly.
Linda had stuffed him down her blouse so his now-cold-blooded majesty wouldn't freeze in the mountain air.
It did explain the indignancy, perhaps, but Linda had bigger problems on her mind.
The hag was gaining on her.

Linda leant right forwards and threw the stick into a dive.
She picked up speed as she shed altitude, but the hag quickly followed suit.
Her feet brushed the snow as she skimmed down the far side of the mountain.
He eyes scanned the slope.
It was a little unfamiliar with this much snow on it; she normally only practiced up here in the summer.
But she soon spotted the stand of trees she was looking for, and the ravine just behind it.

She banked left, sweeping a wide arc across the slope, then dived again, aiming for the ravine.
She clipped the bristles on the ledge as she dropped in, the jerk prompting a panicked croak from her blouse.
She leant forward, nearly hugging the stick, keeping her arms and legs pulled in tight.
It was horribly narrow in here, barely two arm-spans across and mostly sharp turns.
She couldn't normally manage more than half speed through it.
She tried for three-quarters this time.
A broken cackle echoed behind her as the crone followed.

Linda wasn't the best flyer in the world, she knew that.
She wasn't even the best in the kingdom; young Wren up Salwich way could fly circles around her.
And this hag, too, was clearly better than her.
She hadn't even slowed down through the ravine, swerving expertly around the cliffs and gaining quickly.
But the problem with being the best was that there were some things you didn't actually get to practice that much.
Some things that Linda, who was the first to admit that her reach often exceeded her grasp, got to practice all too often.

Linda normally flew this downslope, but she'd pulled in going upstream this time.
The crone was barely a broom-length behind when she found out why.
The two witches rounded a corner to come face-to-face with a cliff, icicles hanging over the lip where the waterfall fell in spring.
Linda grinned and gritted her teeth as she hauled back on the stick, pulling it totally vertical.
One stuck-out leg prevented her from scraping against the cliff, and she kicked herself away from it.
The hag pulled into the climb far more smoothly, pulling almost alongside Linda as she did so.
She grabbed the knife from her mouth and reached out.

Linda swerved away as the knife flashed, tearing her cloak.
It was barely wide enough here for two to fly abreast, nevermind to avoiding getting knifed.
And the sticks were losing speed as they climbed; they weren't supposed to go straight up like this.
The hag soon had her crowded right into the cliff face.
Just one moment longer{\dots}
The knife flashed again.

The sticks stalled.
Horror crossed the crone's face, and she dropped the knife as she clutched at her stick.
Linda kicked out, catching the crone in the ribs and knocking her away from the broom.
Her hands wrestled with her own broom, flipping it over and shoving it back between her legs.
She'd practiced this far too much, although normally with a much greater distance to fall, just in case{\dots}
She willed the broom to restart, and the magic caught.
The ground was rushing up to meet her and she hauled back, wrenching herself out of the dive.
Her feet skidded over the frozen river as she levelled out, just in time, and pointed her stick back down the ravine.
\emph{Fwoomph}.
The hag plunged into a snowdrift beside her.

\storybreak

Linda pulled her stick up outside the castle and dropped neatly to the ground.
She dusted the snowflake off her cloak, blew into her hands a couple of times to warm them up, and fished the king out of her blouse.
She held him up in front of her face.
``Ohh, don't give me that look!
We're getting you up the Queen.
\emph{She} can give you a kiss.''
