\chapter{Golemancy}
\chaplabel{golemancy}

\section{Animating a Golem}

A golem must be animated as part of creation, and the witch doing the animation must be involved in its creation, even if she is not the primary craftswoman.
To animate a golem, a witch simply touches it and wills it life; many consider Golemancy to be a particularly specialised application of \discref{willing}.
A novice golemancer has only enough animating force to maintain one golem at a time.
If she animates a new golem, the previous golem immediately becomes inanimate.
A witch may also withdraw the animating force from a golem she has animated at any time, though if this is to be done urgently (perhaps the golem has gone rogue), the GM may require a Test.

The crafting and animation, although strongly interlinked, are separate processes.
Tests related to the craftsmanship use \skillref{crafting} and an appropriate attribute.
Tests related to the golem's animation, such as giving it instructions, use \testtype{wit}{golemancy}.
A witch can only animate a particular material into a golem if she has taken the appropriate feat.

\section{A Golem's Instructions}

A witch just beginning to dabble in \discref{golemancy} only has the skill to make very simple-minded, single-purpose creatures, although she will learn more nuance later.
These golems are imbued with a single instruction at the moment of their creation.
They will follow this until its completion, whereupon they will simply stand still and await destruction.
The instruction must be very simple, and the golem has minimal ability to improvise around it.
It should not have any conditional aspects, and the golem is unable to respond to any form of communication.
Example instruction are given below.

\begin{itemize}
	\item Deliver this note to the castle.
	\item Fetch my broom.
	\item Kill that man.
	\item Sweep the floor every evening.
	\item Extinguish any fires you see.
\end{itemize}

Additional information necessary to the completion of the task, such as the location of the castle or the identity of an intended victim, may be imparted with the instruction.
The golem will trust this information and cannot adapt if it is wrong, for example if the victim has been disguised.

Giving instructions with nuance, or instructions with multiple linked parts (such as ``go to the castle and kill the King'') requires a Test, with a TN set by the GM based on the complexity of the instruction.
A failure either prevents the golem from animating or, at the GM's option, corrupts the instructions.

\section{A Golem's Statistics}

A golem's physical statistics are determined by the material and method of its construction, and are specified in the appropriate feat.
These include its \attref{might}, \attref{grace}, Speed, Resilience and Shock Threshold.
A golem whose Shock Threshold is met or exceeded by a Damage Test is immediately destroyed, instead of knocked out.
The GM is also advised to apply common sense to other consequences of a golem's construction: for example, a clay golem will sink in water, a wooden golem will float, and a gingerbread golem will go soggy and fall apart.
A golem lacks the other four attributes (\attref{wit}, \attref{will}, \attref{charm} and \attref{presence}) entirely; it automatically fails any Test that would require them.
It has no ranks in any skills.

Initially, a witch can only animate small golems: about a handspan in height.
She doesn't have enough animating force to manage anything bigger, and anything smaller can't hold the magic required.

A golem knows no languages; it cannot read, write, or comprehend speech.
It cannot speak, and furthermore cannot vocalise in any fashion.
The sounds it can make are limited to such things as clapping its hands and stamping its feet.

A golem has senses as good as a human, although only if its craftsmanship gives it the appropriate anatomy.
For example, a gingerbread golem with two currants for eyes can see, but if baked without the currants it will be blind.
A clay golem can only smell if a nose is sculpted upon its face.

\section{Feats}

\feat{Gingerbread Man}{gingerbread-golem}{20}{
	None
}{
	The simplest golems are not baked of clay, but of dough.
	When you bake a humanoid figure from gingerbread, you may animate it as a gingerbread golem.
	The entire preparation and baking process takes approximately half an hour, although an entire batch of golems can be crafted at once if the size of the oven allows.
	
	\materials{Flour, Sugar, Eggs, Butter, Ginger (\herbref{3})}
	
	A gingerbread golem has \negative{2} \attref{might}, 1 \attref{grace} and 15 Speed.
	It has an effective Shock Threshold of 1; it is destroyed if it takes any damage.
	
	Additionally, a gingerbread golem has a limited lifespan.
	After about a week, it grows stale and can no longer move.
	Moisture or water, even a couple of minutes in rain, will destroy it sooner.
}

\feat{Wood Golem}{wood-golem}{15}{
	\featref{gingerbread-golem}
}{
	Wood offers a far more robust and permanent golem than gingerbread.
	When you whittle or assemble a humanoid figure from wood, you may animate it as a golem.
	The time required to do this depends on the size of the golem and the piece of wood you are crafting from.
	Whittling a small golem from an approximately man-shaped piece of wood may take as little as ten minutes, but carving one from a solid chunk of log might take an hour or more.
	Carving a much larger one could take days, and it would likely be faster to assemble it from multiple pieces of wood.
	
	A wooden golem has \negative{1} \attref{might}, 1 \attref{grace} and 10 Speed.
	It has 4 Resilience and a Shock Threshold of 14.
	Damage to a wooden golem cannot be repaired.
}

\feat{Wax Golem}{wax-golem}{15}{
	\featref{gingerbread-golem}
}{
	Wax isn't quite so robust as wood, but it can be very quick to mould and repair.
	When you mould or cast a humanoid figure from tallow or beeswax, you may animate it as a golem.
	The wax or tallow needs to be warmed to be moulded.
	Leaving it in the sun on a warm summer's day provides about the temperature required, as does sitting near a fire.
	Once warmed, the golem can be moulded by hand in a couple of minutes.
	
	A wax golem has \negative{1} \attref{might}, 1 \attref{grace} and 10 Speed.
	It has 2 Resilience and a Shock Threshold of 12.
	Damage to the golem can be easily repaired by the application of a little more warm wax.
	
	Wax golems are susceptible to heat.
	A hot summer's day won't hurt, just make them a little softer, but coming too close to a fire or forge will leave them in a dribbly pool on the ground.
}

\feat{Golem Programming}{change-golem-instructions}{20}{
	\featref{gingerbread-golem}
}{
	You can change the instruction imbued into one of your golems, allowing you to reuse the same golem for multiple tasks.
	Reprogramming a golem requires you to be touching it, and takes a minute of concentration.
	The normal restrictions apply to the new instruction, and the golem can only have one instruction at a time; adding a new instruction removes the previous one.
	You may only reprogram golems powered by your own animating force.
}

\feat{Delegated Programming}{change-golem-instructions-familiar}{10}{
	\featref{change-golem-instructions}
}{
	You've taught your familiar a few tricks of golemancy, and it may use the bond it shares with you to tap into your own animating force.
	Your familiar may reprogram golems, using the same rules as \featref{change-golem-instructions}.
	It reprograms your golems, however, instead of its own.
	If imbuing the new insruction requires a Test, your familiar uses its own \testtype{wit}{golemancy}, not yours.
}

\feat{Dwarf Golem}{medium-golem}{20}{
	\featref{wood-golem} or \featref{wax-golem}
}{
	You can muster enough animating force to create larger golems, about mid-thigh height.
	Such golems have their \attref{might} and Resilience increased by 1, and their \attref{grace} decreased by 1.
	
	Obviously, crafting the body of a larger golem takes more material and normally more time.
	Acquiring an oven, kiln or forge large enough can also present an obstacle for some kinds of golem.
}

\feat{Man-Sized Golem}{large-golem}{25}{
	\skillref[1]{golemancy},
	\featref{medium-golem}
}{
	It takes a lot of clay to make a golem the size of a man, but this pales in comparison to the force required to bring such a golem to life.
	You should know, because you can finally muster that much force.
	Such golems have their \attref{might} and Resilience increased by 2, and their \attref{grace} decreased by 1.
}

\feat{Advanced Instructions}{advanced-golem-instructions}{20}{
	\skillref[1]{golemancy},
	\featref{gingerbread-golem}
}{
	You can imbue your golems with more advanced instructions.
	The instructions can contain several steps, and conditional elements.
	On the whole, the golem can contain instructions that would take no more than a minute to convey by reasonably-paced speech.
	
	The golem still has next to no ability to improvise around the instructions, and will unreasoningly attempt to carry them out until it completes them or is destroyed.
	It can respond to some degree of communication, such as pointing and nodding, if explicitly instructed to.
	However, it still cannot \emph{comprehend} the communication.
	For example, it can follow an instruction to ``go where this man points,'' but not to ``follow this man's directions''.
}
