\chapter{Headology}
\chaplabel{headology}

\section{Feats}

\feat{Curse}{curse}{15}{
	None
}{
	It's a well known fact that someone who believes they will fail is more likely to do so.
	It doesn't take a drop of magic to make that true, but not everyone knows how to leverage it.
	You do.
	
	If someone believes that you have cursed them, or even if you can convince them that they have been cursed by something else, they suffer bad luck.
	Whenever they make a Test, dice that roll a 3 count towards a critical failure.
	The GM is also encouraged to make their critical failures a little more dire.
	This bad luck persists as long as the supposed curse is present in their minds; it might help to remind them now and again.
	
	This only applies if they believe they are under a rather broad curse, or specifically a bad luck curse.
	An overly specific curse---for example, ``May your crops wither in your fields,'' or ``May your nose fall from your face''---does nothing to focus their mind on their own failure and will have no effect.
}

\feat{Mind over Magic}{foil-magic}{15}{
	None
}{
	For all the magic circles and burning incense, magic ultimately comes from the mind.
	Not only do you know this, but you know \emph{how to exploit it}.
	
	If you can convince a practitioner of magic that their magic won't work, then it won't.
}

\feat{Doubt \& Despair}{foil-magic-2}{25}{
	\featref{foil-magic}
}{
	Under your tender care, even the smallest seed of doubt can flourish into a blossoming tree of failure.
	
	If you can make a practitioner so much as doubt the efficacy of their magic, or their own ability to work it, then the magic will either fail to work or, at the GM's option, backfire.
}

\feat{Mind Like a Razor}{headology-weapons-improvise}{10}{
	\featref{willing-tools-improvise}
}{
	If you can convince your foes that what you wield is a weapon, their flesh will believe you.
	You may treat an item you wield or throw as a Knife, Hand Weapon or Thrown Weapon (depending on its size and whether you're throwing it) if you can convince the target that it can cut (or otherwise deal damage) like one.
	A demonstration against an inanimate object, or another foe, will often suffice.
	Even your bare hands can cut like Knives if you convince your foes that they can.
}

\feat{Change Blindness}{headology-stealth}{10}{
	None
}{
	You may hide in plain sight by leveraging the fact that people don't \emph{expect} to see you there.
	This uses a \testtype{charm}{stealth} Test.
	You must remain silent and quite still, though you may creep around very slowly.
	
	In order to make use of this feat, anyone you are hiding from must have no reason to expect to see you, or anyone.
	If they see much out of place---a drawer opened or a vase knocked over---they might look for whoever did it and will immediately spot you.
	Furthermore, you can only use it if the people you are hiding from have some degree of familiarity with the location; they must have seen it before, at least.
	Somebody entering a room for the very first time doesn't know what to expect and will see it as it is, you included.
	
	Lastly, somebody seeing a group or crowd of people has no reason not to expect other people with them.
	This feat does not allow you to hide in such a situation, unless everyone in the group has the feat.
	%TODO: Is there a feat that helps blending in with a crowd?
}
