\chapter{Tools of the Craft}
\chaplabel{equipment}

\section{The Hat}

\section{Broomsticks}

\section{Improvised Tools}

\section{Weapons}

Weapons are divided into several broad categories.
Players are free to describe their character's weapons how they wish, within the bounds of reason, placing them in one of the categories.
Anything a character might find at hand and hit people with can also be placed into a category.

A weapon's accuracy is added a flat bonus to rolls to hit, in place of an attribute.
A weapon's damage determines the number of dice rolled upon hitting.
The highest 3 dice are kept, as always, but the number of dice rolled are determined by the weapon instead of the wielder's skill.
The wielder's Might is added to the damage roll for melee or thrown weapons, but not for bows.

\begin{center}
	\begin{tabu}{X[2.4]XXX[1.2]}
		\toprule
		Weapon & Accuracy & Damage & Range (metres)\\
		\midrule
		Fist & +2 & 3 & Melee\\
		Club & +4 & 4 & Melee\\
		Knife & +2 & 5 & Melee\\
		Hand Weapon & +4 & 5 & Melee\\
		Thrown Rock & +0 & 3 & 5$\times$Might\\
		Thrown Weapon & +0 & 3 & 5$\times$Might\\
		Bow & +0 & 5 & 100\\
		\bottomrule
	\end{tabu}
\end{center}

\subsubsection{Fist}
A punch, a kick, or a headbutt.
Covers any attack you make without any weapon at all.

\subsubsection{Club}
A club, a walking stick, a chair, or a cauldron.
A club is just about anything you pick up and hit someone with.

\subsubsection{Knife}
A knife or dagger.
Easily concealed, and a staple of blood witches.
The short blade costs the wielder reach, but can do as much damage as a sword if you get the enemy in the tender parts.

\subsubsection{Hand Weapon}
A sword, an axe, a mace, a spear, a pike.
This category covers most things actually designed as a weapon and larger than a knife.

\subsubsection{Thrown Rock}
A genuine rock, but also a teapot, a boot or a frog.
Anything you might pick up and throw.
This includes weapons that aren't designed to be thrown.

\subsubsection{Thrown Weapon}
A spear, a knife, a hatchet.
Any weapon you can throw that was actually designed for the purpose.
Rocks from slingshots fall in this category too.

\subsection{Bow}
A bow and arrow.
Also covers crossbows, if the setting includes them.
