Elle Weerstrom looked up from her parsley patch as the air swooshed overhead.
Black fabric flapped.

``Evenin' Linda.
Didn't expect to see you today.
What brings you up 'ere?''

A navy-lined cloak fluttered as the younger witch pulled her broomstick short and dropped to the ground.
``It's young Barnie, Elle.
He's got a mob together, marching on Buckle Hollow.
Says Musgrave's been sleeping with his wife.''

Elle brushed her gloves together, knocking dirt onto the lawn.
``Well, has he?''

``No!
I mean, they might've kissed a bit but{\dots}
They've got torches, Elle!
Pitchforks and torches!
C'mon, grab your broom.
We've got to stop them.''

Elle looked down at the ground, then up at the sky.
She sighed.
``Alright, we'll go.
But we're walkin'; there's a storm brewing.''

Linda looked up.
A single wisp of cloud drifted lazily across the azure sky.
``Looks alright to me.''

``It's on its way, mark my words.
Wouldn't want to be flyin' home in it.''
Elle strode towards her cottage.
``I'm goin' to get my coat.''

\storybreak

Sure enough, the sky was grey when the mob got to Buckle Hollow.
A fine drizzle filled the air.
The farm gate stood open, a figure between the posts in its place.
Her parka was pulled up against the rain, pointed hat tall above her crown.
The mob stopped in its tracks as a crack of lightning cast her silhouette upon them.

``Fine weather for arson, innit?''
Her voice seemed to carry further than it should in the damp air, reaching the ears of all present.
They shuffled their feet in the thickening mud.
``Yer a disappointment, the lot o' yer.''
More feet shuffled.
A voice rose in dissent, but Elle continued over it.

``Now, I know Musgrave ain't the finest man you've all met.
An' I ain't quite sure what he's been up to that's got you all riled up.
But I \emph{am} sure that it ain't nothin' worse than half o' you've done to yer own wives!
Honestly, torches lads?''
The rain intensified and the torches guttered.
One spluttered out.
``What were you goin' to burn?
The barn?
His house?
\emph{Him}?
Put 'em away, men.''

There was another shuffling of feet, and a few torches wobbled noncommittally.
A sudden gust of wind drove the rain sideways for a moment.
Every torch went out with a pathetic cough.
``Get home to yer own wives, an' stop worryin' about other people's.''

With a quiet mumble, a general grumble and a mutter of ``Soddin' linen's gonna be soaked\dots'' the mob turned around and began to trudge the other way.

``An' Barnie!''
The mob stopped in its tracks again.
One man turned around, a few others craned their necks to see.
``She mightn't be kissin' other blokes if you spent as much time in yer own bed as in the gutter out back o' the Head.''
A muffled chuckle ran through the mob before another peal of thunder cut it short.
Collectively, they slank off through the mud.
