\storychapter{Cleaning Up the Blood}

\dropcap{The} barest sliver of moon utterly failed to light the village beneath.
This did not trouble the man who strolled down its streets, heels clicking on the cobblestones.
The blackness rather matched his cloak as it rustled in the breeze behind him.
And his boots, trousers, and cloak.
Only his shirt---white, crisp, recently starched---stood out.
There was blood on it.
It wasn't his.

%He'd killed them tonight.
%He preferred not to do that; it wasn't really sustainable.
%But it had been more than a week since his last excursion, and he'd been so{\dots} hungry{\dots}
%He was better now.
%He'd had his fill.
%It was time to go home, if he wanted to get back by sunrise.
%It was such a long way back to his castle; that was why he didn't come down here as often as he'd like.

The streets ran red, too.
It was something of a waste, he mused.
But he didn't mind.
He'd had his fill already, had it hours ago.
And there would be other villages, for next time.
%For now, however, it was time to go home.
%He'd had to come quite a way to find this village, and it was just as far back.

He passed the inn.
No laughter echoed from within, like it had when he'd arrived.
One of the shutters hung wide open, but no light spilled out.
The barkeep hung over the windowsill.
His face was deathly pale, his eyes blankly staring.
The man strolled on.

Ahead, a young girl rounded the corner, walking into his path.
She wore a grey nightgown and a pointy nightcap, and pushed a broom in front of her.
The man smiled.
Perhaps the night's fun was not yet over.

``Shouldn't you be in bed, little miss?''

The girl looked up.
Her face was untroubled, seemingly oblivious to what had been going on in her village.
``Shouldn't you, too?''

``Ohh, I'm sure my parents won't be missing me.
But---''

The girl looked him up and down before cutting him off.
``No, I don't imagine they would be.''

The smile dropped from the man's face, replaced with a look of annoyance.
``Why don't you run home to your mummy and daddy?'' he snipped.

``They're dead,'' replied the girl.
Her face remained a mask of utter unconcern.
``But I imagine you knew that.''

``Who are you?''

``I could ask you the same question.''

``Who am I?''
The man drew himself up to his full height.
``I am---''

``Kurt.''
The girl cut him off again.
``Spawn of Skew, spawn of Varian.''

The man smiled again.
``Ahh, I see.
You've done your research.
A Hunter, at your age?
I'm impressed.''

``Spawn of \emph{me}.''
The girl smiled at last, showing her teeth.
Her long, pointed teeth.

The man's smile fled, and he died with horror plain upon his face.
