\chapter{Blood Magic}
\chaplabel{blood-magic}

\section{Feats}
\unsetfeatdefault{discipline}
\unsetfeatdefault{tags}

\feat{Blood Binding}{symlink-blood}[sympathetic-magic]{15}{
	\skillref[1]{sympathetic-magic},
	\featref{symlink-taglock}
}{
	\capital{\materialref{blood}} is the most powerful \materialref{taglock}.
	With enough of it, you can forge an even stronger {\symlink}.
	
	If you use \materialref{blood} as the \materialref{taglock} for \featref{symlink-taglock}, it establishes a {\standardsymlink}, instead of a {\weaksymlink}.
}

\feat{Total Bloodbind}{symlink-blood-2}[sympathetic-magic]{10}{
	\skillref[1]{sympathetic-magic},
	\featref{symlink-blood}
}{
	A drop of \materialref{blood} is potent.
	\emph{Soaking} the {\symbol} in \materialref{blood} is simply overwhelming.
	If you soak the whole {\symbol} in the target's \materialref{blood} while using \featref{symlink-blood}, it establishes a {\strongsymlink}.
	Soaking a typical hand-sized \materialref{poppet} or \materialref{effigy} requires 1 dose of \materialref{blood}, but larger {\symbols} might require more.
}

\feat{Blood Thicker Than Water}{blood-combine}[brewing]{10}{
	\noprereq
}{
	\capital{\materialref{blood}} runs in the family.
	You may create a person's \materialref{blood} by combining the \materialref{blood} of their ancestors in appropriate proportions.
	In the simplest form, you simply mix together the \materialref{blood} of both their parents.
	However, you may do this recursively: for example, you may form the target's \materialref{blood} from the \materialref{blood} of their mother and both their paternal grandparents in the correct ratio.
	
	The resulting \materialref{blood} is treated as having come from the person for all purposes.
	It is now a \materialref{taglock} connected to them, and no longer to any of the ancestors.
	Volume is conserved, and thus this produces a total amount of \materialref{blood} equal to the total amount combined, which may be less than a dose.
	The \materialref{blood} may be mixed in any vessel, and doesn't require heating.
}

\feat{Bones of the Ancestors}{blood-combine-2}[brewing]{10}{
	\skillref[1]{brewing},
	\featref{blood-combine}
}{
	Using \materialref{blood} from a person's ancestors is all well and good while they're alive, but much trickier once they're dead.
	You've learned a useful workaround, as long as you can dig up up their bones or ashes.
	You may use \materialrefplural{taglock} other than \materialref{blood} when you use \featref{blood-combine}, ignoring ratios, as long as you still use \materialref{blood} from at least one ancestor.
	
	Doing things this way is more complicated, and thus requires a heated \mixcreationref{cauldron}.
	The amount of \materialref{blood} produced is equal to the amount of \materialref{blood} used, ignoring all other \materialrefplural{taglock}.
}
