\discipline{Brewing}{brewing}{Brewer}{Brewers}

\dropcapdiscref{brewing} may well be the least magical of witchcraft's disciplines.
In fact, it is not even restricted to witches.
Many wise folk can cook up a remedy for the most common ailments.
And almost every village has someone to brew their hops into beer.

As such, the feats in this discipline are not necessary for common brewing; they represent only those brews that require some trick, skill, or guarded piece of knowledge.
Anyone, even those without ranks in the \skillref{brewing} skill, may attempt to create more common brews.
Some, such as alcohol, are relatively simple.
A Test may determine the quality of the result, but almost anyone can create \emph{something} appropriate.

Remedies for diseases and ailments are a little more complicated.
That's not to say that it takes much skill to boil up a few strips of willow bark, but half of the common knowledge of folk medicine is ungrounded superstition.
It requires a \testtype{ken}{brewing} Test to create an appropriate remedy for most ailments, with the {\tn} of the Test determined by the rarity and severity of the ailment.
This normally requires access to a wide array of \materialrefplural{herb} and a boiling cauldron; the {\tn} might be increased as normal by trying to make do with inadequate materials.

For particularly simple remedies---such as helping a person to sleep, or easing an ache---it may suffice to use the raw \materialref{herb}, without any \discref{brewing}.
In this case, a \testtype{ken}{botany} Test might suffice.

\section{Brew-Toting}

A \practitioner{brewing} is not just a witch who can brew potions, but also a witch who tends to carry them.
It is assumed that a \practitioner{brewing} carries at least one dose of each potion she knows how to make; normally several doses of each.
Unless she begins providing potions to groups much larger than her coven, or keeping people dosed on a particular potion for hours on end, she should have plenty of doses.
If she does do this, however, the GM is welcome to declare that she runs out and has to brew some more.

This assumes, however, that she has ready access to all the ingredients required to brew her potions.
In most cases, this simply means that the witch has access to her {\garden}, and sufficient \skillref{botany} skill that her {\garden} contains the required \materialrefplural{herb}.
If she doesn't have ready access to the required ingredients, she will either have to gain access somehow, or improvise replacements.
In either case, the GM is free to have these endeavours roleplayed, and require Tests for them.

\section{Brews for Everyone}

Most \practitioners{brewing} sample their own fare.
But the main benefit of \discref{brewing} comes from sharing one's potions.
The other members of your coven; their familiars, and yours; friendly village folk; even an \featref{animal-companion}; everyone can benefit from your brews---except golems and the undead.

As such, many \practitioners{brewing} fall into something of a supportive role within their coven.
So don't be afraid to share your potions, to allow everyone to do what they do even better.

This doesn't have to come at the cost of doing your own thing, however.
You can easily brew a dose or two of every potion for every member of your coven, and can even make the other witches carry their own doses.
The brunt of your \discref{brewing} work will be done during downtime, leaving you plenty of time to get involved when you're out and about.

\section{Creation and Application}

The \skillref{brewing} skill and \discref{brewing} discipline don't just cover potions brewed in a cauldron.
Potions, poultices, poisons, tinctures, salves, ointments, even beer, mead, wine and spirits.
Witches have many ways of turning \materialrefplural{herb}, and even other ingredients, into more useful forms.

Each feat that allows a witch to prepare such a mixture lists the method of preparation and delivery.
The rules of such methods are presented below.

\subsection{Brewing and Chewing}

Different methods of preparation require different equipment, and take different periods of time.

\mixcreation{Cauldron}{cauldron}{
	Most potions are brewed in cauldrons, filled with water and brought to boil.
	This requires, obviously, a cauldron, as well as a fire to heat it.
	A smaller kettle might do in a pinch, but requires a Test.
	A full cauldron will typically yield several doses.
	Brewing in a cauldron requires around 15 minutes to bring the water to the boil, and another minute to mix the potion.
}

\mixcreation{Poultice}{poultice}{
	A poultice doesn't need to be brewed at all; the ingredients are simply chewed into a paste.
	Some of the more dangerous poultices should definitely be ground with a mortar and pestle, however, rather than allowed anywhere near the mouth.
	Creating a poultice requires less than a minute.
}

\mixcreation{Still}{still}{
	Some potions, or spirits, need to be distilled.
	This requires quite a lot of dedicated equipment, a carefully maintained heat source, and several hours.
}

\subsection{Method of Delivery}

Although most potions are drunk, there are many ways to get a mixture into a person's body.
Some faster, some slower, and some far easier to inflict on an unwilling victim.

\mixdelivery{Drink}{drink}{
	A \mixdeliveryref{drink} is about half a litre of liquid that must be drunk to take effect.
	It can be quaffed as an {\action}, and takes effect after 1 {\round}, unless specified otherwise.
}

\mixdelivery{Spike}{spike}{
	A \mixdeliveryref{spike} is a much smaller quantity of liquid than a \mixdeliveryref{drink}, little enough that it could be slipped into a glass of wine without being obvious.
	It can be drunk willingly, but typically isn't.
	It takes effect after 1 {\round}, unless specified otherwise.
	
	In a glass of wine or a cup of water, one dose will normally go unnoticed until it is too late.
	A second dose causes a noticeable change in taste, scent, or colour, which will typically be noticed unless the drinker isn't paying much attention.
	Larger and stronger drinks can conceal more doses, however.
}

\mixdelivery{Topical}{topical}{
	A \mixdeliveryref{topical} mixture is applied to the skin.
	It typically requires more than an {\action} to smear it on, or bind a wad in place.
	It generally only takes effect after a few minutes, but can kick in a little faster if applied to a wound or a mucous membrane.
	Some need to be applied to the correct part of the body.
	For example, if it is to treat a wound, it should be applied to the wound, and if it is to enhance the eyesight, it should be applied to the eyes.
}

\mixdelivery{Snuff}{snuff}{
	These mixtures are boiled or ground down to a powder, which must be inhaled into the nostrils.
	They can be taken as an {\action}, and take effect immediately.
	Giving them to someone unwilling requires forcing them to inhale in some fashion, but it can usually be achieved if you have the target at your mercy for a minute or more.
}

\mixdelivery{Injury}{injury}{
	These mixtures, typically harmful ones, must be delivered into the bloodstream via an injury.
	The most expedient way to do this is to smear it on an arrow, a \weaponref{blowgun} dart, or an edged weapon, requiring an {\action}.
	It's good for one cut, but otherwise remains on the weapon until rubbed off or washed away.
	Beware rain.
	It takes effect immediately, unless specified otherwise.
}

%TODO: Gaseous? Needs to be stored in an air-tight bottle?
%TODO: Incense? Needs to be burned, evaporated?

\subsection{Time to Effect}

Mixtures applied by different methods typically require different lengths of time to take effect, as specified above.
Some mixtures require more or less time to take effect, as specified in their descriptions.
If quaffed on a creature's {\turn}, a mixture that takes effect after 1 {\round} comes into play at the start of their {\turn} in the next {\round}.

The GM may allow a character to make \attref{might} Tests to stave off the effects of an unwanted mixture.
This might extend the time to effect by two or three times, but should not allow them to avoid the effect entirely, except in the most exceptional cases.

\subsection{Vomiting}

For ingested mixtures (a \mixdeliveryref{drink} or \mixdeliveryref{spike}), vomiting before it takes effect can massively ease this \attref{might} Test, and even allow avoiding the effect entirely.
Ingesting an emetic herb, such as \herb{veratrum or the toadstool ``emetic russula''}{2}, serves to induce vomiting.
However, these herbs do not act quickly enough to help with a mixture that takes effect in just 1 {\round}.
To prevent these, a witch might look into \featref{vomit-potion}.

\section{Side-Effects}

While some of the more noxious mixtures a witch can brew produce adverse effects by design, these are not the only ways a potion can hurt.
Many mixtures come with adverse side effects all by themselves, and these are compounded by the dangers of overdosing and combining brews.

\subsection{Overdosing}

Many potions carry harmful effects when taking too many doses.
These typically only occur if multiple doses would be in effect simultaneously; taking another dose after the first has worn off is safe unless specified.
Some of the effects of overdoses are given explicitly, but many are given as general guidelines.
The GM is left to adjudicate in the latter case.
Typically the worst effects of overdosing can be staved off with a \attref{might} Test, with the {\tn} affected by how many excess doses have been taken, and how close in succession they were taken.

If you wish to keep a brew's effect going continuously, and you have access to additional doses of the brew, it is easy enough to take another dose as you feel the first wearing off.
You can do this without any risk of overdosing.

\subsection{Mixing Mixtures}

Mixing multiple potions can have adverse and unexpected effects.
These kick in when a character is under the effect of two substances that both affect the same attribute, or other statistic.
The effects are unpredictable.
The GM is free to apply any appropriate penalty, possibly calling for a \attref{might} Test to avoid or alleviate the effects.
The following table is provided for inspiration.
The GM may roll 2 six-sided dice and compare their sum against the table to randomly determine an effect, if desired.

\begin{simpletable}{rX}
	\toprule
	\dice{2} & Effect\\
	\midrule
	2 & Apply severe overdose effects of the first mixture.\\
	3 & Exhaustion, unconsciousness and/or oxygen deprivation.\\
	4 & Apply moderate overdose effects of the first mixture.\\
	5 & Ignore any positive effects of the first mixture.\\
	6 & Double any detrimental effects of the first mixture.\\
	7 & Re-roll twice on the table, taking both results.\\
	8 & Double any detrimental effects of the second mixture.\\
	9 & Ignore any positive effects of the second mixture.\\
	10 & Apply moderate overdose effects of the second mixture.\\
	11 & Twitching, seizure, overheating and/or organ failure.\\
	12 & Apply severe overdose effects of the second mixture.\\
	\bottomrule
\end{simpletable}

The effect of painkillers---to ignore {\damage}---does not count as a statistic for the purpose of mixing substances.
As such, a character may safely be under the effects of multiple painkillers as long as their other effects do not overlap.
Additionally, withdrawal effects (such as those of \featref{mental-stimulant}) do not count for mixing.
%TODO: Use a better example when I have a withdrawal effect that modifies statistics.

\section{Antidotes}
\seclabel{antidotes}

While vomiting can help to avoid the effects of a potion \emph{before} they kick in, ending the effects once they are ongoing requires an {\antidote}.
\capital{\antidotes} exist not just for poisons, but for any potion or other mixture with an ongoing effect.
Each {\antidote} will only work on potions it is designed to counteract, however.
It will specify in its description what sorts of potion it is effective against.

\capital{\antidotes} can be delivered through different methods, just like potions themselves, and take the usual length of time to kick in.
Once it kicks in, the {\antidote} ends all ongoing effects of any potions it is designed to counteract.
One dose of {\antidote} can end the effect of several different potions, as long as it is designed to counteract all of them, or several doses of the same potion.

When an {\antidote} takes effect, treat it as though the duration of the potion has expired.
An {\antidote} ends both beneficial and detrimental effects, but can only end \emph{ongoing} effects.
Any {\damage} that has been dealt stays dealt.
Overdose effects are ended, but organs that have failed are not repaired, and so on.
Withdrawal effects (such as \featref{mental-stimulant}), are not ended by an {\antidote}.
In fact, taking an {\antidote} causes the withdrawal effects to kick in, as it ends the effect of the potion.

The effect of an {\antidote} is immediate, not ongoing.
This means that you cannot have an {\antidote} to an {\antidote}.

\section{Feats}

\feat{Numbing Painkiller}{painkiller-grace}{15}{
	\noprereq
}{
	\mix{cauldron}{drink}{\Herb{willow bark}{2}}
	%Willow bark contains natural NSAIDs.
	
	The drinker may ignore 1 point of {\damage} for a few hours, but loses 1 \attref{grace} for the same duration.
	Two doses may be effective simultaneously.
	Further doses cause paralysis, and possibly organ failure.
}

\feat{Dimming Painkiller}{painkiller-mental}{10}{
	\noprereq
}{
	\mix{cauldron}{drink}{\Herb{poppy seed}{2}}
	%Poppies contain natural opioids.
	
	The drinker may ignore 1 point of {\damage} for a few hours, but loses 1 \attref{ken} and \attref{wit} for the same duration.
	Two doses may be effective simultaneously.
	Further doses cause unconsciousness, and possibly cessation of breathing.
}

\feat{Blurring Painkiller}{painkiller-heed}{15}{
	\noprereq
}{
	\mix{cauldron}{drink}{\Herb{barley}{2}}
	%Barley contains a lot of phenols, which are used in the synthesis of a lot of pharmaceuticals.
	%Malt barley is also used in producing alcohol, which can justify the blindness.
	
	The drinker may ignore 1 point of {\damage} for a few hours, but loses 1 \attref{heed} for the same duration.
	Two doses may be effective simultaneously.
	Further doses cause blindness, which can become permanent.
}

\feat{The Hard Stuff}{might-potion}{20}{
	\noprereq
}{
	You know how to make a drink that'll really put hairs on a man's chest.
	Or a woman's, at that.
	
	\mix{still}{drink}{Alcohol, \herb{apple}{2}}
	%Scumble is 'mostly apples'.
	
	The drinker gains 1 \attref{might} for a few hours, and loses 2 \attref{wit} and \attref{heed} for the same duration.
	A second dose will render the drinker unconscious.
	Further doses are dangerously poisonous, causing vomiting, seizures and oxygen deprivation.
}

\feat{The Pure}{might-potion-2}{15}{
	\skillref[1]{brewing},
	\featref{might-potion}
}{
	\mix{still}{drink}{Alcohol, \herb{fennel}{2}, \herb[aniseed]{green anise}{3}, \herb{wormwood}{3}}
	%Wormwood, sweet fennel and green anise are the three herbs in absinthe.
	
	You brew your drink clean and pure.
	This functions as \featref{might-potion}, except it decreases \attref{wit} and \attref{heed} by only 1 point.
}

\feat{The Green Fairy}{might-potion-3}{20}{
	\skillref[2]{brewing},
	\featref{might-potion-2}
}{
	\mix{still}{drink}{Alcohol, \herb{fennel}{2}, \herb[aniseed]{green anise}{3}, \herbcreature{grand-wormwood}{5}}
	
	Your drink gives men the strength of horses.
	You don't want to see what it does to horses.
	This functions as \featref{might-potion-2}, except it increases \attref{might} by 2 points.
}

\feat{Empathogen}{charm-potion}{20}{
	\noprereq
}{
	\mix{cauldron}{drink}{\Herb{violet}{2}}
	%Violet contains piperonal, a common precursor to the empathogen MDMA (ecstasy).
	%Violets were also emblematic flowers of Aphrodite and Priapus.
	
	This potion affords the drinker a greater sense of empathy and connection with those around them.
	The drinker gains 1 \attref{charm} for a few hours, and loses 2 \attref{will} for the same duration.
	A second dose causes agitation and paranoia, instead reducing \attref{charm} by 1.
	Further doses cause the drinker to overheat, suffering heat stroke, and may lead to internal bleeding and organ failure.
}

\feat{Alpha's Potion}{presence-potion}{20}{
	\noprereq
}{
	\mix{cauldron}{drink}{\Herb{onion}{2}}
	%Onions have an obvious and pervasive smell that makes people cry.
	
	This potion grants the drinker increased confidence, a slightly louder voice, and a certain indefinable \emph{obviousness}.
	There's a certain smell goes with it, a sort of threat pheromone, but it sits below the conscious level for all but the most attentive people.
	The increased confidence that the potion provokes tends to go a bit too far, however, veering into arrogance.
	If the drinker isn't careful, they come across as, frankly, a right prick.
	
	The drinker gains 1 \attref{presence} for a few hours, and loses 2 \attref{charm} for the same duration.
	A second dose causes a total loss of social graces, decreasing \attref{charm} by a further 2 points, without further increasing \attref{presence}.
	It also causes the victim to sweat profusely and smell strongly of onions.
	Further doses cause degeneration into raving, incoherent lunacy, and turns the sweat into a glistening mucus that coats the skin.
}

\feat{Stimulant}{mental-stimulant}{25}{
	\noprereq
}{
	\mix{cauldron}{drink}{Ants, vinegar}
	%Ants have a gland which can produce phenylacetic acid.
	%This can be converted to phenylacetone using acetic anhydride, which can be derived from acetic acid.
	%Acetic acid is the main component of vinegar (beside water). Vinegar is simply fermented from ethanol.
	%Phenylacetone can be converted to amphetamine using the Leuckart reaction and formic acid (also from ants).
	%Amphetamine is a widely-known nootropic (cognitive enhancer).
	
	The drinker gains 1 \attref{wit}, \attref{will} and \attref{heed} for about an hour.
	The potion also staves off tiredness for the duration. %TODO: A mechanical effect for this?
	After the potion wears off, you pay the price of your temporarily enhanced performance.
	You suffer a \negative{1} penalty to all rolls for the next 24 hours.
	Additional doses within this period are ineffective.
	
	Being under the effect of two doses simultaneously causes a headache that counteracts the increased attributes.
	Further doses can cause bleeding into the brain and death.
}

\feat{Stimulant Dragging}{mental-stimulant-extension}{15}{
	\skillref[1]{brewing},
	\featref{mental-stimulant}
}{
	A slight change to the formula of your \featref{mental-stimulant} allows its effect to be extended by additional doses.
	Drinking another dose as one begins wearing off extends the effect and staves off the withdrawal.
	However, the body can only sustain such enhanced performance for so long.
	Whenever you take a dose after the first, make a \attref{might} Test.
	The {\tn} is 9 for the second dose, and increases by 3 for every subsequent dose.
	On a failure, you pass out, gain no benefit from the additional dose, and the withdrawal effects kick in.
	You cannot be roused for several minutes.
}

\feat{Hysterical Strength}{physical-stimulant}{15}{
	\noprereq
}{
	A person's muscles are stronger than they normally get to use, strong enough to break their own bones.
	There's a good reason you don't get to use the full strength, you see.
	But you've figured out how to unlock that extra potential, and damn the consequences!
	
	\mix{cauldron}{drink}{\Herb{joint pine}{3}}
	%Joint pine is Ephedra, which contains ephedrine.
	%Ephedrine is similar in structure and function to epinephrine (adrenaline).
	%Adrenaline is oft-blamed for hysterical strength.
	
	The drinker gains 1 \attref{might} and 1 \attref{grace} for a few minutes.
	For the duration, any strenuous activity causes the drinker to suffer a \dice{2} {\damagetest}.
	Strenuous activity includes the \actionref{dash} and \actionref{attack} {\actions}, any Test using \attref{might} or \attref{grace} (excluding {\damagetests} as part of the \actionref{attack} {\action}), and other things at the GM's discretion.
	
	Being under the effect of two doses simultaneously does increase \attref{might} and \attref{grace} further, but causes {\damagetests} as a result of any movement at all; only lying still is safe.
	Further doses cause seizures, triggering the {\damagetests} themselves.
}

\feat{Hysterical Restraint}{physical-stimulant-2}{20}{
	\skillref[1]{brewing},
	\featref{physical-stimulant}
}{
	A slight refinement to the formula for your \featref{physical-stimulant} affords the drinker more control over their newfound strength, limiting the harm they do to themselves.
	The {\damagetests} caused by your \featref{physical-stimulant} are only \dice{1}.
}

\feat{Hysterical Empowerment}{physical-stimulant-3}{20}{
	\skillref[1]{brewing},
	\featref{physical-stimulant-2}
}{
	By concentrating your \featref{physical-stimulant}, you can make it stronger, but more dangerous.
	Whenever you brew \featref{physical-stimulant}, select a number from 1 to 3.
	The potion provides this number as a bonus to both \attref{might} and \attref{grace}, but also determines the number of dice rolled on the {\damagetests} it causes.
	
	Doses beyond the first only ever provide 1 additional point to the attributes, regardless of their strength.
	However, they still carry all the same drawbacks.
}

\feat{Hysterical Overload}{physical-stimulant-4}{10}{
	\skillref[2]{brewing},
	\featref{physical-stimulant-3}
}{
	You can give your \featref{physical-stimulant} potion \emph{even more power}!
	When you use \featref{physical-stimulant-3}, you can brew potions that give an attribute bonus up to \positive{5}, although this raises the number of dice rolled for the {\damagetests} as normal.
}

\feat{Hysterical Moderation}{physical-stimulant-mitigate}{20}{
	\skillref[2]{brewing},
	\featref{physical-stimulant-3}
}{
	You can fine-tune the low end on your \featref{physical-stimulant} potions.
	When you use \featref{physical-stimulant-3} to brew a potion that provides exactly \positive{2} to \attref{might} and \attref{grace}, the {\damagetests} it causes are only \dice{1}.
}

\feat{Brew of Claws}{weapon-potion}{15}{
	\noprereq
}{
	\mix{poultice}{topical}{\Herb{bramble stems}{1}}
	
	Smeared onto hands or paws, this mixture causes the digits to calcify, and harden into wicked claws.
	They return to normal after about an hour.
	Until then the creature rolls additional dice for {\unarmed} {\damagetests} using the clawed appendages.
	
	A creature without an effective attack gains one, and rolls 2 dice, while a creature with an existing attack rolls at least 3 dice.
	A human, or other creature with proper hands, rolls 5 dice.
	
	However, the hardening makes it difficult to use the hand normally.
	You roll 1 die less on Test that use the clawed hand's manual dexterity, including attacks with weapons held in that hand.
	You may apply this \mixcreationref{poultice} to only one hand, granting you the improved {\damagetest} while leaving the other hand unaffected.
	If used on on a paw or other appendage of a non-human, the GM is free to apply penalties appropriate to the afflicted appendage.
	
	This \mixcreationref{poultice} uses only the stems of the brambles, not the thorns, so it is safe to chew.
	It leaves the mouth a little hardened, but causes no major problems.
}

\feat{Oakenhide Brew}{resilience-potion}{15}{
	\noprereq
}{
	\mix{cauldron}{drink}{\Herb{parsnip}{2}, \herb{oak bark}{2}}
	%Parsnip stems and leaves can cause skin rash.
	
	This potion causes the drinker's skin to harden in patches, sprouting bark-like growths.
	It lends a remarkable toughness, at the price of an awful stiffness.
	
	It takes effect over the course several minutes, eventually increasing the drinker's \statref{resilience} by 1, but reducing their \attref{grace} by 2.
	A second dose further increases \statref{resilience}, but makes the drinker so stiff that they cannot move at all.
	Further doses risk making this immobility permanent.
	
	The growths flake off after a few hours, returning the drinker to normal.
	The energy expended in growing and shedding these patches takes a couple of extra meals to recoup---someone using this potion more than once a day for an extended period simply can't keep up, and will starve.
}

\feat{Stonehide Brew}{resilience-potion-2}{15}{
	\skillref[1]{brewing},
	\featref{resilience-potion}
}{
	You've found a rare succulent plant which disguises itself as a pebble.
	Boiling its tough skin, you can create a potion that lends the hardness of stone to those who drink it.
	
	\mix{cauldron}{drink}{\Herb{parsnip}{2}, \herb{pebble plant}{4}}
	%Pebble plants are the genus Lithops.
	
	This functions as \featref{resilience-potion}, except it increases \statref{resilience} by 2.
}

\feat{Ironhide Brew}{resilience-potion-3}{15}{
	\skillref[2]{brewing},
	\featref{resilience-potion-2}
}{
	Using the legendary strength of the \creatureref{ironwood} tree, you can grow a skin that will nearly turn knives.
	
	\mix{cauldron}{drink}{\Herb{parsnip}{2}, \herbcreature{ironwood}{5}}
	
	This functions as \featref{resilience-potion}, except it increases \statref{resilience} by 3.
}

\feat{Supple Hide}{resilience-potion-mitigate}{15}{
	\skillref[2]{brewing},
	\featref{resilience-potion}
}{
	The growths produced by your \featref{resilience-potion}, \featref{resilience-potion-2}, and \featref{resilience-potion-3} are more supple, less restrictive.
	They reduce \attref{grace} by only 1.
}

\feat{Rapid Hide}{resilience-potion-fast}{15}{
	\skillref[1]{brewing},
	\featref{resilience-potion}
}{
	Your \featref{resilience-potion}, \featref{resilience-potion-2}, and \featref{resilience-potion-3} are more potent, faster-acting.
	They take effect in only 1 round.
}

\feat{Hedgehog Hide}{resilience-potion-spines}{10}{
	\featref{resilience-potion},
	\featref{weapon-potion}
}{
	When you brew a \featref{resilience-potion}, \featref{resilience-potion-2}, or \featref{resilience-potion-3}, you may cause the growths to form spikes.
	You must add a hedgehog's spine to the potion when you brew it.
	
	The spines grant the affected creature the same bonuses to {\unarmed} {\damagetests} as the \featref{weapon-potion}.
	The does not suffer the penalty to manual dexterity caused by the \featref{weapon-potion}, only the penalty to \attref{grace} caused by the \featref{resilience-potion}.
	
	Furthermore, when a creature with these spines is hit by an {\unarmed} attack, the attacker suffers a \dice{2} {\damagetest}.
}

\feat{Berserker Broth}{rage-potion}{15}{
	\noprereq
}{
	\mix{cauldron}{drink}{\Herb{rose}{2}}
	%Roses are blood-red and thorny.
	
	The potion fills the drinker with an all-consuming, empowering rage.
	For a few minutes, they gain 1 \attref{might} and \attref{will}.
	However, for the same duration, they want nothing more than to murder everything; people and other creatures.
	
	A person under the effect of this brew retains all their skills and knowledge; they will still use weapons and tools, and a witch retains her full ability with magic.
	However, the rage is urgent, and overwhelming.
	The drinker will not plan or delay their murders; they would rather draw a weapon and charge in than laying an ambush.
	They still have a sense of self-preservation, but it is a second priority, behind murder.
	Thankfully, however, the potion does not instill any desire to extend its own effect, even if doing so would give more time to murder people.
	
	If a player character drinks this, the GM \emph{may} allow them to keep control of their character, as long as they act in a appropriate fashion.
	However, the GM reserves the right to take control of the character if she is being insufficiently murderous, or insufficiently indiscriminate.
	
	A second does further increases \attref{might} and \attref{will}, but leaves the victim so mad that they cannot do anything but curl into the foetal position and scream.
	Further doses risk making this madness permanent.
}

\feat{Brighteye Drops}{sight-potion-light}{15}{
	\noprereq
}{
	\mix{cauldron}{topical}{\Herb{eyebright}{2}}
	
	Dripped into the eyes, this concoction enhances the eyesight.
	However, the user's night vision is utterly shot.
	
	For the next hour, the user rolls an extra die on \skillref{perception} Tests relying on sight in well-lit places.
	This replaces any existing bonus dice from an ability such as a \familiarrefpossessive{raptor}, but applies on top of the normal dice granted by the \skillref{perception} skill.
	
	However, the user is almost utterly blind in dimly lit locations.
	Most inside locations will be dimly lit, unless they have large or numerous windows, or an abundance of lamps.
	
	Further doses produce no additional effect.
	If taken alongside \featref{sight-potion-dark}, the user is blinded in both light and darkness.
}

\feat{Darkeye Drops}{sight-potion-dark}{10}{
	\noprereq
}{
	\mix{cauldron}{topical}{\Herb{carrot}{2}}
	
	The opposite of the \featref{sight-potion-light}, this potion grants excellent night vision, at the cost of an over-sensitivity to light.
	For an hour after application, the user suffers no penalties in low-light conditions, though they are as blind as anyone in complete darkness.
	
	For the same duration, bright lights---such as sunlight, or a lantern flame---are blinding.
	Looking into such light blinds the user for about a minute.
	
	Further doses produce no additional effect.
	If taken alongside \featref{sight-potion-light}, the user is blinded in both darkness and light.
}

\feat{Twitching Eyes}{sight-potion-switch}{15}{
	\skillref[1]{brewing},
	\featref{sight-potion-light},
	\featref{sight-potion-dark}
}{
	Normally, taking \featref{sight-potion-light} and \featref{sight-potion-dark} at the same time leaves the user completely blind.
	However, a slight change to the brewing process allows each to act as an {\antidote} to the other, in addition to its normal effects.
	This allows you switch back and forth between the two effects, through repeated applications of the mixtures.
}

\feat{Molenose Powder}{smell-potion}{10}{
	\noprereq
}{
	\mix{cauldron}{snuff}{\Herb{jasmine}{3}, an animal's nose}
	%Jasmine is noted for its fragrance.
	%The flowers look star-shaped, like a star-nosed mole's nose.
	
	Snorted into the nostrils, this concoction delivers an overwhelming fragrance of jasmine, which quickly fades to leave the user's sense of smell greatly enhanced.
	This enhanced sense lasts a few minutes, however, the concoction strikes the user completely blind for the same duration.
	
	Their sense of smell becomes fine enough to allow them detect people moving around a room, in real time.
	This should make walking around at a slow pace relatively easy, despite the blindness.
	With a difficult Test, they might even be able to run, without running into anything.
	Successfully attacking people, or evading attacks, is all but out of the question.
	
	For the duration, the user rolls 2 extra dice on \skillref{perception} Tests relying on smell.
	This replaces any existing bonus dice from an ability such as a \familiarrefpossessive{rat}, but applies on top of the normal dice granted by the \skillref{perception} skill.
	
	Additional doses burn out the nostrils, leaving the user with no sense of smell \emph{or} sight.
}

\feat{Mole's Eyes}{smell-potion-2}{10}{
	\skillref[1]{brewing},
	\featref{smell-potion}
}{
	By cutting \featref{smell-potion} with \herb{eyebright}{2}, you can ameliorate its effect on the eyesight.
	
	\mix{cauldron}{snuff}{\Herb{jasmine}{3}, \herb{eyebright}{2}, an animal's nose}
	
	This functions as \featref{smell-potion}, except it doesn't render the user completely blind.
	It still has a drastic effect on their vision, leaving even nearby objects as little more than blurry shapes.
	However, the user can safely walk around without trouble, and can even run without \emph{much} risk of running into something.
	Attacking or evading is even possible, albeit still at a large penalty.
	They also halve the number of dice they roll on \skillref{perception} Tests relying on sight.
}

\feat{Recovering Mole}{smell-potion-3}{15}{
	\skillref[2]{brewing},
	\featref{smell-potion-2}
}{
	A subtle change to the formula of your \featref{smell-potion} makes it last much longer, without hindering the eyes any more.
	
	The duration of your \featref{smell-potion} (and \featref{smell-potion-2} concoction) increases to an hour.
	However, this only increases the duration of its effect on smell; vision still recovers to normal after only a few minutes.
}

\feat{Hot Cuppa}{warm-potion}{10}{
	\noprereq
}{
	A cup of hot tea can certainly warm the body briefly, but this concoction can keep you warm for \emph{hours}.
	
	\mix{cauldron}{drink}{\Herb{horseradish}{2}}
	
	For a few hours, this concoction keeps the drinker warm and comfortable even in chilly weather.
	They can even go out in the snow without wrapping up, though freezing temperatures are enough to cause discomfort.
	They are highly resistant to hypothermia and frostbite---only an extended period in icy water is typically sufficient to afflict them.
	
	In warm temperatures, however---such as a midsummer's day, or near a large fire---the drinker is more susceptible to overheating, and heat stroke.
	Being under the effect of two doses at once similarly causes overheating, leading to heat stroke, and organ failure with further doses.
	If taken alongside \featref{cold-potion}, the drinker's temperature fluctuates wildly, switching between the two effects randomly.
}

\feat{Iced Cuppa}{cold-potion}{10}{
	\noprereq
}{
	Fed up of sweltering over a cauldron's fire, you've brewed up a solution.
	
	\mix{cauldron}{drink}{\Herb{mint}{2}}
	
	For a few hours, this concoction keeps the drinker comfortably cool even in sweltering heat.
	They won't suffer from heat exhaustion, or heat stroke.
	However, they are not protected from {\fire}, or anything else that would cause burning.
	
	However, the drinker becomes more susceptible to cold temperatures.
	Air temperatures below freezing are dangerous.
	Similarly, being under the effect of two doses at once causes hypothermia, and further doses risk frostbite in the extremities.
	If taken alongside \featref{warm-potion}---or any brew that causes overheating---the drinker's temperature fluctuates wildly, switching between the two effects randomly.
}

\feat{Flaming Phlegm}{fire-attack-potion}{15}{
	\featref{warm-potion}
}{
	\mix{cauldron}{drink}{\Herb{chilli pepper}{3}}
	
	This concoction mixes with the drinker's saliva to form a volatile substance that ignites when spat.
	The effect lasts for a few minutes after drinking the potion.
	
	The drinker---as long as they are a creature with at least 1 die on {\unarmed} {\damagetests}---can spit of gob of flaming phlegm using the \actionref{attack} {\action}.
	This attack has a \positive{0} accuracy bonus, and a range of 2 metres.
	On a hit, it causes no {\damagetest}, but {\ignites} the target at \dice{2}.
	
	Drinking this potion has no immediate drawbacks, although flammable saliva does carry obvious own risks.
	The drinker is advised to keep their mouth shut when they are not using it, to minimise the chance of accidental ignition.
	Drinking multiple doses simply increases the risk of spontaneous oral combustion.
}

\feat{Vomiting Drops}{vomit-potion}{15}{
	\noprereq
}{
	\mix{cauldron}{spike}{\Herb{mistletoe}{2}}
	
	These drops cause immediate vomiting.
	They take effect immediately, and act quickly enough to be effective against ingested mixtures that take effect in just 1 {\round}.
	The drinker loses at least one {\action}, and usually several, while they heave and retch.
}

\feat{Sleeping Solution}{sleep-potion}{10}{
	\noprereq
}{
	\mix{cauldron}{spike}{\Herb{camomile}{2}}
	
	This soporific kicks in quite slowly, but the drinker should be asleep with 10 minutes, and remain that way for a few hours.
	A few doses are quite safe, and will accelerate the effect a little, but too many can put the victim into a coma.
}

\feat{Garlic Solution}{garlic-potion}{15}{
	\noprereq
}{
	\mix{still}{spike}{\Herb{garlic}{2}}
	
	A distillation of purest garlic essence, this solution goes straight to the sinuses and burns something awful.
	It takes effect immediately.
	The drinker must make a {\tn} 12 \attref{might} Test or pass out immediately, unable to be roused for several minutes.
	For the next hour or so, they have no useful sense of smell and a full breath to the face robs others of their own sense of smell.
	They can be easily tracked by scent as the stuff leaks from their pores.
	Remarkably, however, the solution itself has no obvious scent until ingested.
	
	Vampires suffer far worse, of course.
}

\feat{Projection Potion}{projection-potion}{15}{
	\skillref[1]{brewing},
	\featref{garlic-potion},
	\featref{projection-start}
}{
	\mix{still}{spike}{\Herb{garlic}{2}, \herb{morning glory}{3}}
	%Morning Glory is an Aztec entheogen.
	
	This solution has enough kick to knock the mind right out of the body.
	It takes effect immediately, ejecting the drinker's mind into the {\mentalrealm}.
}

\feat{Projection Poison}{projection-poison}{20}{
	\skillref[2]{brewing},
	\featref{projection-potion}
}{
	You can brew a variant of \featref{projection-potion} that is delivered through the bloodstream, rather than the mouth.
	This makes it a little easier to inflict on an unwilling target.
	
	\mix{still}{injury}{\Herb{garlic}{2}, \herbcreature{moly}{5}}
	
	This functions just as \featref{projection-potion}, except for its different method of delivery.
	Bear in mind that the victim can very quickly re-enter their own body.
	However, it will probably have fallen over before they do so, and a prepared mind could jump in and {\possess} the body first.
}

\feat{Medium's Potion}{medium-potion}{10}{
	\skillref[1]{brewing},
	\featref{projection-potion},
	\featref{medium}
}{
	By brewing the herb used to enter a \featref{medium} trance into the right concoction, you can skip the tedious meditating.
	Or feed it to someone else, if you like.
	
	You may include \herb[stinking nightshade]{black henbane}{2} in the recipe of your \featref{projection-potion}, or your \featref{projection-poison}, if you can make the latter.
	If you do so, then the user's body is placed into a \featref{medium} trance as their mind is ejected.
	
	As normal, the \featref{medium} may end the trance at any point by re-entering her body.
	However, a soul ready and waiting to {\possess} the body still gets the first chance.
}

\feat{Nettle's Bite}{poison-pain}{15}{
	\noprereq
}{
	You can grind up nettles while preserving, and enhancing, their sting.
	The resulting mixture won't do much lasting harm, but it hurts like nobody's business.
	
	\mix{poultice}{injury}{\Herb{nettle}{1}}
	
	If this poison is delivered by an attack, the target's \statref{shock-threshold} is treated as being 1 lower against the {\damagetest}.
	This does not increase the {\damage} dealt, but is more likely to send the target into {\shock}.
	
	Preparing this \mixcreationref{poultice} by chewing is incredibly painful, and leaves a swollen tongue that makes speaking difficult for hours---a mortar and pestle are recommended.
}

\feat{Festering Poison}{poison-infection}{10}{
	\noprereq
}{
	\mix{poultice}{injury}{\Herb{bloodwort}{2}}
	
	This wicked concoction invariably causes bleeding and infection when applied to a wound.
	Cleaning out the infection requires medical attention, and typically a Test.
	\capital{\damage} caused by a weapon coated in this poison, or any wounds it is applied to, will not heal until the infection has been cleared.
	If the infection goes untreated for several days, it can become lethal.
	Even if the infection is treated and the wound heals, it will often leave a wicked scar.
	
	Chewing up one or two doses of this \mixcreationref{poultice} is relatively safe, as long as you have no wounds around your mouth.
	Chewing a larger batch can cause bleeding and subsequent infection in the mouth.
}

\feat{Embalming Fluid}{embalming-fluid}{10}{
	\skillref[1]{brewing}
}{
	\mix{still}{drink}{Alcohol, ants, iron shavings}
	%Basing this on formaldehyde.
	%Formaldehyde is produced industrially by oxidation of methanol.
	%One catalyst used for this is a mixture of metals including iron.
	%Ants are for the connection between formic acid and formaldehyde.
	
	This fluid is not intended for consumption, but rather for soaking corpses.
	When used in {\embalming}, the corpse can preserved for years, or even decades if it is well cared for.
	One dose suffices to {\embalm} a rat or bird, but two doses are required for a cat, or more than a dozen for a human.
	
	If, for some reason, the potion is imbibed, it proves quite toxic.
	It causes pain, nausea and convulsions, progressing, over the course of a few minutes to permanent blindness, and probably death.
}

\feat{Healing Salves}{brewing-healing}{10}{
	\skillref[1]{brewing}
}{
	You know a wide range of minor poultices, salves and remedies for cuts, bruises and other physical injuries.
	As long as you have access to a reasonable supply of various \herbtypeplural{2}, and time to chew up poultices, you may use your \skillref{brewing} skill in place of your \skillref{healing} skill on Tests to heal people and creatures of most physical injuries.
	Setting broken bones and performing surgery still requires \skillref{healing}.
	
	Similarly, you may use your \skillref{brewing} rank in place of your \skillref{healing} rank when determining the {\damage} healed by a patient during a {\dayofrest}.
}

\feat{Healing Stimulant}{brewing-healing-2}{10}{
	\skillref[1]{brewing},
	\featref{brewing-healing}
}{
	\mix{cauldron}{drink}{\Herb{goldenrod}{2}}
	%Goldenrod is used as traditional medicine to heal wounds.
	
	This potion stimulates the body's natural growth and repair.
	However, it only works as long as the drinker rests.
	With one dose of the potion, the drinker heals twice as much damage during a {\dayofrest}.
	That is, they heal 4 points of {\damage}, plus any healed by their physician; the physician's healing is not doubled.
	The drinker needs to eat at least twice as much as normal, to fuel the regeneration.
	
	Taking more than one dose in a day causes overgrowth; scabs and scars develop even where the drinker is unwounded, and they can become deformed.
	Healing is no faster, but they need to eat even more.
}

\feat{Life Juice}{brewing-healing-3}{20}{
	\skillref[2]{brewing},
	\featref{brewing-healing-2}
}{
	\mix{cauldron}{drink}{\herbcreature{arbor-vitae}{5}}
	
	This potion acts like \featref{brewing-healing-2}, but much faster, and without the need to rest.
	The drinker heals up to 4 points of {\damage} over the course of a few minutes.
	However, they suffer one level of {\exhaustion}, affecting \attref{might} and \attref{grace}.
	
	Two simultaneously effective doses cause the same effect as an overdose of \featref{brewing-healing-2}.
	Additionally, consecutive doses continue to cause more {\exhaustion}, which may cause the drinker to pass out, as usual.
}

\feat{Regrowth Potion}{brewing-healing-regrow}{10}{
	\skillref[2]{brewing},
	\featref{brewing-healing-2}
}{
	Salamanders have the incredible ability to regrow lost limbs.
	By adding a salamander's leg to the recipe for your \featref{brewing-healing-2} or \featref{brewing-healing-3}, you can grant the same ability to the drinker.
	
	The drinker can regrow missing body parts.
	Fingers and ears will grow back with a single dose, while a limb or eye takes several doses.
	A dose of \featref{brewing-healing-2} must be accompanied by a {\dayofrest}, as usual, and multiple doses must be spread across multiple days.
}

\feat{Hair of the Dog}{brewing-healing-taglock}{20}{
	\skillref[2]{brewing},
	\featref{brewing-healing}
}{
	Folk medicine claims that, when bitten by a dog, you should apply a few hairs of the dog that bit you to the wound.
	It never worked with the raw hairs, but you've discovered a quick preparation that fixes that.
	
	\mix{poultice}{topical}{A \materialref{taglock} from the creature that attacked you, \herb{dock leaves}{1}, alcohol}
	
	Applied to a wound, this mixture will completely heal the wound over the course of a few minutes.
	This allows you to heal all the {\damage} dealt by the creature whose \materialref{taglock} is used in creating the poultice.
	However, it can only heal {\damage} dealt by {\unarmed} attacks---punches, bites, claws, and the like---not any {\damage} dealt with a weapon.
	Furthermore, it will only close wounds and heal {\damage}; it cannot regrow missing parts.
}

\feat{Exposure}{poison-resistance}{15}{
	\skillref[1]{brewing}
}{
	You've tasted a few too many of your own concoctions, but you're still alive.
	In fact, you're beginning to build up a bit of resistance.
	
	You roll an extra die on Tests to resist poisons or the like.
}

\feat{Bottled Sobriety}{antidote-alcohol}{10}{
	\skillref[1]{brewing}
}{
	\mix{cauldron}{spike}{\Herb{tea leaves}{4}}
	%Tea contains caffeine, commonly touted as sobriety-inducing.
	
	This concoction acts as an {\antidote} to alcohol of any kind, as well as any other mixture which lists alcohol as an ingredient, such as \featref{might-potion}.
	Alas, it won't cure a hangover.
}

\feat{Universal Antidote}{antidote-all}{25}{
	\skillref[3]{brewing},
	\featref{antidote-alcohol}
}{
	You have achieved one of the philosophers' stones of \discref{brewing}, the universal {\antidote}.
	
	\mix{cauldron}{snuff}{A toad's brain, \herbcreature{moly}{5}}
	%Moly is a mythical herb from the Odyssey that gave immunity to sorcery.
	%The toad brain is firstly a reference to toadstones, supposedly an antidote to any poison.
	%Secondly, it is established by familiars that toad's have a knowledge of brewing.
	%This knowledge might include the knowledge to counteract any brew.
	
	Taken by itself, this powder is inert; it has no effect.
	However, when this powder is combined with any other mixture, that mixture loses its effect and becomes precisely its own {\antidote}.
	This requires one dose of \featref{antidote-all} and one dose of the other mixture.
	The application method of the other mixture is not changed by this process: a \mixdeliveryref{spike} remains a \mixdeliveryref{spike}, \mixdeliveryref{snuff} remains \mixdeliveryref{snuff}, and so on.
}
