\storychapter{Whodunnit}

\dropcap[``]{Whodunnit}, huh?
We're good at these, eh Faith?''
Old Granny Ada grinned at the corpse, while Faith held her nose against the stench.

``I would have thought they'd have learned by now,''
replied Faith, words slightly distorted by her pinched nose.
``You were banging on young Gregory's door within an hour of him pinching that candlestick, and when Percy killed that cow{\dots}''

``Aye, aye, but this fella's bin clever, see?
Carter here's bin dead three, four days, by the smell.''
Granny tapped her nose and gestured at Faith, who rolled her eyes.
``Well, I certainly ain't gonna sit here three days, looking that far back.
Can't stay awake that long at my age, lemme tell you!
How about you?''

Faith looked pensive for a moment, then sighed.
``Well what are we going to do, then?''

``A bit of good old-fashioned sleuthing, that's what.
We don't have to use magic for everything, you know!
You and Scarlett go have a look-see round the village;
I'll poke about here.''

\newcommand\foretellingstoryrepeat{%
	``Alright, I'll see what I can find.''
	Faith strode outside the woodsman's cottage and cast her gaze around.
	Scarlett, her magnificent peregrine falcon, was preening on top of the thatch.
	She put two fingers in her mouth,
}%
{\foretellingstoryrepeat}
and whistled sharply.
Scarlett swooped down and alighted on her thick, leather glove.
Faith smiled at her, and the two strode down towards the village.

\storybreak

It was mid-afternoon, several hours later, before Faith finally caught up with the murderer.
Turned out Carter had been involved in a bit of smuggling, riding weapons over the border to Houghton under the dead of night.
John, the miller, and Abel, the smith, had been in on it too.
But when Carter wanted out, he'd taken all the gold off the last run, and threatened to rat.
Abel had been happy enough to turn King's evidence, when he found out Carter was dead.
So now she was hot on John's tail.

He wasn't coming quietly, though.
He was doing a runner for the stables; if he got a horse, he'd be in Houghton before Faith could make it back to the cottage for her broom.
Scarlett was high overhead, keeping an eye on him, while Faith sprinted after.
He was faster, though, and fitter.
He pulled ahead.

Scarlett screeched as John reached the stables, and dived.
Faith burst through the doors behind him, scanning the building.
No sign of him.
He must have gone into one of the stalls, but which one?
He'd be bridling the horse right now{\dots}

Faith stopped, blinked hard, took one step backwards, and opened her eyes.
Her Hindsight came on.
Time began playing out, backwards, before her eyes.
The door behind her slammed closed again, as she saw a version of herself run backwards through it.
A few second later---or rather, earlier---John ran, backwards, out of the third stall on the left.

Faith smiled, and blinked again.
Time jumped back to normal; she'd only lost a few short seconds, and had time to catch her breath while she watched.
She ran left, turning into the stall.
John stood there, nearly as out of breath as she was, bridle halfway onto the horse.

In his other hand was a knife.

Faith stopped in horror.
She stepped back, snatching a riding crop off the wall to defend herself.
John lunged.
Pain blossomed in Faith's chest.
She coughed, spitting blood.
Then his elbow caught her in the face, and everything went black.

\storybreak

{\foretellingstoryrepeat}
and---

A sharp whistle cut through the air, from the cottage.
Faith, confused, turned and headed back inside.
She reeled backwards as the door swung open; the smell had somehow got immeasurably worse.
The rank stench of faeces and the sharp tang of blood had joined the putrid odour of decay.
Fearing the worst, Faith steeled herself and burst inside.

Carter's corpse was split right open, a clean slice from crotch to chest.
His entrails spilled across the ground.
Granny Ada stood over him, a knife dripping blood in one hand and a solemn look on her face.
Faith's mouth opened and shut like a goldfish.
``Wha---?''

``John done it.
I wasn't gonna interfere; figured you'd learn a lot, sleuthing.
But turns out he kills you after you catch him.''
Granny jerked a thumb at the mirror, standing against the wall.

Faith looked in the mirror.
She saw herself reflected there, but not her as she was right now.
Her reflection was soaked with sweat and smeared with grime; and she was obviously knackered from running.
More importantly, she had a knife sticking out of her chest.
She was lying on the floor, her blood pooling in the straw on the floor.
``But---''

``You found the axe he done it with, buried shallow in his garden.
We can dig that up; ought to be evidence enough.''

``But{\dots}
How did you know?''

``Ohh, all credit to you, lass!
You caught him, or you was gonna.
I just watched you do it.''
Granny chuckled.
``Fine bit of sleuthing, really.
Just a shame about the stabbing.
I'll show you how you done it, later.''

``And{\dots}
And the{\dots}''
Faith gestured incoherently at the entrails on the floor.

``Ohh, them was just to find the moment to watch.
I'll pop them back in, sew him up.''
She looked at Faith's very pale face.
``You don't have to help, don't worry.''

``I think I need to go and lie down{\dots}''
Faith finally summoned a complete sentence.
``Outside{\dots}''

``Alright,''
replied Granny.
``You got five minutes, then we're gonna go have a chat with John.''
