\storychapter{}

\dropcap{The} battle was near wrapped up now; the sun about an hour above the horizon.
The defenders had sallied out with great success, crippling the invaders' catapult before it could be brought to bear.
The invaders were routed now, but casualties had been high on both sides.

Ivy Shotworth bustled back and forth in the castle's foyer.
She'd set up a sort of field hospital; injured men laid out on what rugs were available, or on the bare flagstones.
More were coming in quickly.
Most of them hobbling, some being carried.
One of the maids was helping with triage.
One of the patients---a leg wound that'd be alright in a week, tops---was keeping a couple of irons hot in the fire.
She'd had to use them too often already.
Those were three limbs that their owners wouldn't be seeing again in a hurry.

Ivy hurried to the next patient.
Young Ned, the blacksmith's son, now missing a finger from his left hand.
But he'd been lucky; it was a clean cut, and the next finger was barely nicked.
She sloshed some of the contents of her hip flask over the stump, then bound it tight with a strip of cloth.
That'd do for now, but she'd have to come back to it later.
Hopefully whoever'd kept his blade sharp enough for a cut like that had kept it clean too.

A bark came from behind her as she moved down the line of patients.
Monty was standing by the door, whining quietly and indicating outside with his head.
He was a bloodhound; a big one.
He'd gone out to help find wounded, and it seemed someone had turned up.
Ivy exchanged a quick glance with him, then spun around and shouted down the hall.

``Grover!
Bit of help here!''
The din in the foyer quieted down for a moment as everyone turned to look---patients, soldiers, and maids.
Grover broke off his conversation and hurried over.
She beckoned to indicate he should follow her, then headed outside after Monty.
Grover followed wordlessly.

The pointed hat was useful like that.
No hesitation, no awkward questions.
Sure, she'd earned the respect; cures, midwifing, even surgery.
She made herself more than useful around here.
But the hat was a constant reminder.
It'd make itself useful for what was coming, too.

Monty led them around the back of the castle, where two men lay.
It looked like they'd been caught away from most of the fighting, doing who knows what, when they'd run into one another.
One was definitely dead; he had the other man's sword sticking out both sides of his chest.
The other man---wearing the invaders' colours---had a nasty gash down his flank.
He was lying on the ground, clearly unconscious, but his chest was rising and falling.

Ivy strode to the breathing man's legs and bent down, hooking her arms under his knees.
``You get his shoulders.''

Grover gave her a look.
\emph{Are we really doing this?} it asked.
\emph{Are we really helping the enemy?}

But he didn't actually vocalise it.
It was the hat that did that.
That, and the look she gave him back.
It said, in no uncertain terms, \emph{Yes.}

\storybreak

Between them, they carried the man back into the castle.
Most of the army was in the courtyard now, pulling themselves out of their armour, and cleaning their weapons.
Sir Egland---one of the kingdom's two knights---had just dismounted, and was striding around the courtyard, clapping people on the back, bellowing about a ``jolly good fight'', and ``giving them bastards what for''.
Ivy usually preferred to avoid him; talking to folks like him was always more trouble than it was worth.
But there was no helping it this time.

Eyes turned to follow them as they crossed the courtyard---all these eyes had been on red alert for enemy colours quite recently, and they weren't going to miss them now, even if the man wearing them was out cold.
Again, the hat kept them all silent.
But their attention eventually drew Sir Egland's own eyes; even he couldn't miss two dozen soldiers falling silent and staring in one direction.

``Oi! What the hell do you think you're doing?'' he demanded, striding over to Ivy and Grover.
He caught up with them as they reached the edge of the courtyard.

Ivy met his eyes, but didn't slow down, still carrying the man through the doors, into the foyer.
She kept her voice calm as she replied, betraying neither fear, nor anger.
``I'm taking a wounded man for treatment, Sir.''
Grover faltered a little, nearly letting go of the man's shoulders.
But Ivy shot him a quick glare, and he carried on.

``That's an enemy soldier, you stupid hag!
You don't treat them!''
He spat the word `them'.

``He might have been five minutes ago, Sir.
Now he's a wounded man, same as any other in this hall.
I \emph{will} treat him.''
She and Grover set the man down on the rug nearest the door.
Grover backed off a little way, joining the silent circle of faces that had gathered to watch the spectacle.
Ivy straightened up, staring into Sir Egland's eyes.

She had to look up a bit to do it; he was a good head taller than her.
The point of her hat came a little above his crown, though, and he'd taken his helmet off.
He still had his maille, his gauntlets, his greaves, and his boots on, of course.
And he had to be at least twice her weight even without that.
She stood between him and the wounded man.

``Anyone who's with them''---he spat the word again---``is to be put to death!''
He drew his sword.
The \emph{shiiing} as it slid out of its scabbard echoed around the otherwise silent hall.

``I won't let you do that.
And I advise you to put that away.''
A quiet crackling came from her hand, as she flexed her fingers.

``Why, you---''
Sir Egland's sword swung; a massive overhand chop, straight for her clavicle.

Ivy's hand flashed upwards, with a glint of steel.
The wounded man had carried a dagger, tucked into his boot.
She'd slipped it free as she set him down, and she brought it up now, catching Egland's sword.
\emph{Claang!}

She stepped in and grabbed the hilt of the sword, twisting.
Egland, still surprised to find any resistance, lost his grip.
The sword clattered to the flagstones, and Ivy sent it spinning away with her foot.
Egland watched it go.

He wheeled on her again, face full of rage.
His fists came up, and he swung.

Ivy sidestepped neatly, grabbing hold of the arm as it went past.
Then she stepped in again, twisting.
She hooked her leg around his, pulled him off balance, then threw him over her hip.
He landed on the next rug over from the the man they'd carried in.
Ivy dropped a knee onto his belly, whipped a damp rag out of a pocket, and clamped it over his mouth.
Egland struggled for a moment, then his eyes rolled back.
He went still.

A circle of wide eyes stared at the spectacle; mouths were agape.
The room held its breath.
The silence hung for a moment; even now, the hat commanded silence.
Until, finally, a deep rumbling began.

Sir Egland had started to snore.

Ivy stood up, stuffed the rag back into a pocket, and wiped her hands.
``Alright everybody, back to work!
There're wounds to tend!
I'll look after \emph{him}''---she nearly spat the word, \emph{nearly}---``later.''

\storybreak

Grover plucked up the courage to come over to her, as she dressed the gash on the man they'd carried in.
``I though you was going to turn him into a frog, miss.
Or{\dots} or smite him!''

Ivy's voice was soft when she replied.
``Come now, Grover.
That wouldn't be right.
He's not a bad man, really.
He's just got some{\dots} wrong ideas.
I'll set him straight when he wakes up.''

``And besides,'' she finished, ``I don't hold with magic.''

Grover was shocked.
``But aren't you a witch?''

``I am.
Have you seen anything to suggest otherwise?''

``Well, umm{\dots}''
Grover thought for a while.
``No, I guess{\dots}
But don't witches do magic?''

``Do I look like I need to?''

There was silence for a while, as Grover thought, hard.

``Good.''
Ivy grinned.
``Now, go soak this cloth, would you?''
