\storychapter{The Tower}

\dropcap[``]{Ooh}, she's going to regret this!''
Linda spun on her heel and began pacing the other way down the cell.
``Lock us up in the tower, would she?
Like we're princesses or something?
She's the damn princess!
Not us!''
She shook with anger, tugging on the ropes that bound her hands behind her back.
``Ooh, she'd better just wait until I get my hands on my broom again.
I'm going to fly down there and give her \emph{such} a slap.''

``She'll be Queen by nightfall.
Maybe she thought she could move out of her tower early.''
Mabel chuckled to herself.
She was sitting cross-legged on the flagstones, leaning back against the bars to the next cell.
Her eyes were closed, her face a total mask of unconcern.

``Total cheek, is what it is!
After everything we did for her parents!''
She stalked over to the high, barred window and whistled out through it.

Elle spoke up from where she lurked in the shadows of the corner.
``Pipe down, would yer?
Even if Hector gets yer broom, you ain't gonna be able to fly it in 'ere, an' yer just gonna draw attention if yer keep hootin' like that.''
She sighed.
``Now come on over 'ere and get yer ropes off.''

Linda looked over at Elle, noticing the ropes lying neatly coiled at her feet.
She trudged over and offered up her bound wrists.
Elle's hands, freed of their own bonds, brushed hers.
Linda felt the ropes gently slide away, neatly undo their knots, and saw them coil themselves on the floor in the corner.

``Alright then,'' said Linda.
``No brooms.
You've got a plan then?''

``Me?
No.
Reckon Mabel's got this one.''

``Might do,'' conceded Mabel, but her smile marked her a plain liar.

``Mabel?'' Linda asked.
``But what've we got to draw with in here?''

``Ohh, I've always got something.
Blood.
Splash of water.
I've got a full bladder.''
She grinned unpleasantly.
``But Elle can just scratch this one, if you'd be so kind?''

``Any time,'' replied Elle.

``But it'll have to wait 'til lunch,'' continued Mabel.
``I need the gruel.''

``Lunch?'' exclaimed Linda.
``But they'll have started the coronation by then!''

``Aye, but they won't've finished it,'' said Elle, unmoved.

%``And how do you even know we'll \emph{get} lunch?''

%``They always bring the prisoners lunch, noontime,'' replied Mabel.
%``Even done it myself, couple of times.
%The guards're good folk, and our dear princess won't've told them not to.
%She's been kind of busy, yerknow.''
%Mabel smiled strangely.
%``It'll be gruel, though,'' she muttered.
%``Always is.''

``But can't you just make these stones jump out?'' asked Linda, slapping one of the masonry walls.
``We can climb down the outside of the tower; I've got a trick for---''

``Could.
Won't.''
Elle lay down, settling onto the bare flagstones.
``Wouldn't get us far enough from what Mabel's plannin', I reckon.''

Mabel gave a chuckle of agreement.

\storybreak

Lunch came at noon---gruel, as Mabel has said.
Elle had finally scratched a ritual circle into the floor, using her bare fingernails under Mabel's direction.
The two of them had done something to hide it---even Linda couldn't focus on it properly, and she knew it was there.
The guards hadn't had a clue.

Mabel set her bowl of gruel on the floor, untouched, when the guards left.
Then she grabbed the sleeve of her dress in her teeth, and tore a strip off.
She dangled it in the air.
``If you would, Elle?''

Elle looked up from her deep concentration upon her own bowl of gruel and shot a glance at the strip of fabric.
It smouldered and burst into flames.
Mabel dropped it in the middle of the floor.
A quick kick with her foot undid whatever they'd done to hide the circle, which faded back into Linda's view.

``Now, I'd stand back a bit, if I was you.''
Mabel started a bit of quiet muttering over the circle, and the flame in the centre slowly swelled.
Soon, it had consumed the entire scrap of fabric and was merrily burning away on the flagstones.
Mabel scooped up her bowl of gruel and started heating it over the fire.
Linda watched, waiting for whatever they'd been planning to finally happen.

After a few moments of awkward silence, Mabel began tucking in.
She looked up at Linda as she slurped down the second mouthful.
``You can heat yours too, if you want.
Gruel's no good cold.''

Linda snapped, for the second time that day.
``This was to heat your gruel?'' she screamed.
``We've been sitting here for four hours so you could \emph{heat your gruel?}
Would it really kill you just to \emph{tell me what you're planning?}''

``Alright, alright, no need to yell.''
Mabel scarfed down the last of her gruel.
``The fire's just for a bit of fun, really.
I mean, look around you.
We can step right out of here whenever we want.''

``We{\dots}
What?''
Mabel gave an expansive gesture, and Linda looked around.
``I{\dots}''
She looked back at Mabel, confused.
%``How?''

%Mabel sighed.
%``Alright, look out the window.
%That hill over yonder, to the right a bit.''

%Linda stood on tiptoes to have a look.
%The grassy hill poked out amongst the farms.
%Atop its crown stood an ancient ring of standing stones.
Then realisation dawned.

``You mean the stones, don't you?
But{\dots}
You need a standing stone for that, don't you?
And a---''

Mabel chuckled.
``I said to look around.
The tower's made of stone, ain't it?
And it's certainly still standing.''
The chuckle had a worrying undertone this time.
``And that''---she gestured at the bars that fronted the cell---``is in iron rod.
Which, near as I can tell, is buried right in the ground, eh?''

``But that's \emph{cheating}!''

``Damn right it is.''
Mabel grinned.
``You done with those oats, Elle?''

``Aye, they're done.''
Elle upended her gruel bowl into Mabel's hand, pouring out dried grains without a trace of watery soup.

``Let's be off, then!''
Mabel strode to the cell's bars, and grabbed hold of one.
She muttered under her breath as rattled it.
``Pop on through, quick sharp!''
She stepped back and waved the others towards the bars, tossing the oats over her shoulder, into the fire.

Elle went first, and Linda followed hot on her heels.
There was a lurch, and a sudden sensation of falling, and then she was staggering out into grass.
Mabel caught her shoulder as she came up behind her, twisting her around and pointing into the distance.
``Just watch there.''

Linda looked up, away from the hilltop they stood on, past the circle of stones they'd arrive in.
The castle was just across the way, two tall towers poking up from the corners.
And, as she watched, the near tower erupted.
Masonry launched skyward, a pillar of roaring flame chasing it up.
The sound reached them several seconds later, as they saw the chunks of rock raining down on the roof.

``Reckon that'll put a damper on the coronation, eh?'' asked Mabel.

``I reckon,'' said Elle.

Linda sighed.
