\discipline{Necromancy}{necromancy}{Necromancer}{Necromancers}

\section{Reanimation and Resurrection}

Many \practitioners{necromancy} draw a distinction between reanimation and resurrection.
Reanimation is a crude process, somewhat akin to \discref{golemancy}.
It's nothing more than the application of raw animating force to a corpse, to stand it up and get it shuffling around again.
The creature retains its instincts, its muscle memory and the like, but that's as much through what is left of its biology as it is through the will that animated it.
The results of reanimation are known as the undead.

Resurrection, by contrast, brings a creature back back to life in full, and the creature's {\soul} is restored to its body.
There may be a few ill effects of the process, not to mention whatever killed it in the first place, but these can typically be recovered from.
For all intents and purposes, the creature is just as much alive as it was in the first place.

In theory, at least.

The trouble is that nobody has ever achieved true resurrection.
Dozens of witches and hundreds of charlatans have all claimed to.
Many have even come incredibly close, but there has always been some slight snag.
The search, of course, continues, but many have given up all hope that it is possible.

\section{Reanimation Rituals \& the Undead}
\seclabel{reanimation-rituals}

Undead, the products of reanimation, come in many different forms: \undeadrefplural{zombie}, \undeadrefplural{skeleton}, \undeadrefplural{ghoul}, and more.
The rituals to reanimate these creatures are just as numerous, but they all share mark{\'e}d similarities.
And many \practitioners{necromancy}, who consider these rituals to form the heart of the discipline, learn a lot of techniques for improving them.

For convenience, the feats which modify these rituals refer to them collectively as the {\reanimationrituals}, and the feats which provide them are listed here.
\begin{itemize}
	\item \capital{\featref{animate-zombie}}
	\item \capital{\featref{animate-skeleton}}
	\item \capital{\featref{animate-ghoul}}
	\item \capital{\featref{animate-draugr}}
	\item \capital{\featref{animate-sea-draugr}}
	\item \capital{\featref{animate-fossil}}
	\item \capital{\featref{animate-fire-skeleton}}
	\item \capital{\featref{animate-shade}}
	\item \capital{\featref{animate-wraith}}
\end{itemize}
Obviously, you cannot perform any variant of a ritual unless you have the feat to perform that ritual in the first place.

\subsection{The Lurching Dead}

The statistics and capabilities of an undead creature are based upon the statistics and capabilities of the creature whose corpse it is raised from.
Each variety of undead comes with its own associated changes to these statistics, listed in the following sections.
Many changes, however, are shared between all reanimated creatures, and are listed here.

Most reanimations do nothing to heal {\damage} to the corpse: both {\damage} suffered before and during death, and any further {\damage} done to the corpse since then.
If this reduces its \statref{shock-threshold} to 0 or below, the corpse is too mangled to successfully reanimate.
Some reanimations also require an unrotted corpse.
It usually takes a little over a week before a corpse becomes too rotted for such a reanimation, although temperature and moisture can alter this.
Corpses can be preserved by {\embalming}.

A reanimated creature loses its memory and identity.
It retains general knowledge such as how to hunt, but forgets such information as the location of its den, and loses any mannerisms that distinguished it in life.
It retains {\generalskills}, but loses {\specialityskills}, {\disciplineskills}, and feats.

Most reanimated creatures also lack many biological processes that they had in life.
They do not need to breathe, eat, drink, or sleep.
They are immune to poisons, diseases, and the like.
Additionally, they cannot heal themselves or be healed, and are unaffected by potions and such.
They do not suffer from {\shock}, and are simply deanimated if they suffer a {\damagetest} equalling or exceeding their \statref{shock-threshold}.
They still suffer from {\exhaustion}, although they don't need sleep to recover from it; just rest.
Lastly, they cannot produce any venom or other such substances, so any benefit of a venomous bite, sting, or the like is lost.

When you reanimate a creature, the resulting undead is under your control.
You control it mentally, without need for verbal instructions or gestures.
However, you can only maintain control while you remain conscious, and nearby---at least near enough to see or otherwise sense it.
If you fall asleep, fall unconscious, die, get too far away from the creature, or leave your body for the {\mentalrealm}, you lose control over it.
A novice \practitioner{necromancy} has no way to regain control over an undead creature once she loses it, other than to kill the creature and reanimate the corpse again.

When you lose control over an undead, it regains free will.
It begins to act as an animal of its kind normally would, seeking out whichever food it would normally eat.
However, it is eternally, ravenously, and insatiably hungry.
It's generally considered good practice to put an undead down, rather than to lose control of it.

A witch who can raise an undead creature learns to control one, and only one, at a time.
When you reanimate a second creature, you must either relinquish control of the previous one, or immediately lose control of the new one.

\undead{Zombie}{Zombies}{zombie}{
	A \undeadref{zombie} is the simplest reanimation possible; the corpse, fully clothed in its own flesh, is simply stood up and walked around as it is.
	Is it a clumsy creature, with most of the mind rotted away as well, and it only grows worse as the corpse rots further.
	
	A corpse reanimated as a \undeadref{zombie} loses 2 points from all attributes except \attref{might} and \attref{will}.
	Its \statref{speed} is halved.
	Its \statref{shock-threshold} increases by 2, however.
	If it could fly, it is now too clumsy to do so.
	
	A corpse reanimated as a \undeadref{zombie} is not healed of any {\damage}.
	The reanimation also requires that the corpse is unrotted.
	Furthermore, it does nothing to slow the rot.
	A \undeadref{zombie} that rots too far loses animation.
}

\undead{Ghoul}{Ghouls}{ghoul}{
	A \undeadref{ghoul} retains greater mental and physical faculties than a \undeadref{zombie}, but this comes at a dangerous price.
	A \undeadref{ghoul} is sustained only by consuming the flesh of its own kind.
	For example, a rabbit \undeadref{ghoul} must consume rabbit flesh, and a human \undeadref{ghoul} must consume human flesh.
	
	A corpse reanimated as a \undeadref{ghoul} loses 2 points from its \attref{ken}, \attref{wit}, and \attref{charm} scores.
	It retains the ability to fly, if it could in life.
	
	A corpse reanimated as a \undeadref{ghoul} is not healed of any {\damage}.
	However, an animated \undeadref{ghoul} may heal {\damage} by consuming the flesh of its own kind.
	An entire corpse is sufficient to restore {\damage} equal to its maximum \statref{shock-threshold}, with smaller portions restoring proportionally smaller amounts.
	A ghoul can consume an entire corpse in less than a minute, and there is no end to its hunger: it could consume corpses for hours on end without being sated.
	\undeadrefplural{ghoul} created from partial corpses, such as with \featref{undead-head}, need only eat their own mass to count it as a full corpse.
	
	Reanimating a corpse as a \undeadref{ghoul} also requires that it is unrotted.
	While animated and fed at least one full corpse each week, however, a \undeadref{ghoul} does not continue to rot.
	
	Lastly, a \undeadrefpossessive{ghoul} unnatural hunger makes it harder to control than most undead.
	It must be fed a full corpse at least once a week, or it always breaks free of the \practitionerpossessive{necromancy} control.
	Even a \undeadref{ghoul} reanimated as a \undeadref{souled} is not immune to this: it goes insane with hunger after a week, and does not return to sanity until it has fed again.
}

\undead{Draugr}{Draugar}{draugr}{
	With all the moisture drawn out of a corpse, it is not only prevented from rotting, but can also be freed from the bloated clumsiness that afflicts \undeadrefplural{zombie}.
	The result is a \undeadref{draugr} (plural \undeadrefplural{draugr}).
	The better preservation also grants it a better memory and senses.
	
	A corpse reanimated as a \undeadref{draugr} loses 2 points from its \attref{wit}, \attref{charm}, and \attref{presence} scores.
	It retains the ability to fly, if it could in life.
	
	A corpse reanimated as a \undeadref{draugr} is not healed of any {\damage}.
	The reanimation requires that the corpse is unrotted, and it also must be {\embalmed} by desiccation (drying out).
	This usually is usually done using salt, but can happen naturally to creatures that die in deserts.
	
	To remain animated, the \undeadref{draugr} must be kept dry; water bloats the corpse, starts it rotting again, and immediately ends its animation.
	It might manage a 30 second sprint through light rain, but heavier rain is too much.
	Given a heavy leather coat, it might just about be able to travel through rain, but the moisture will still get to it in a couple of hours.
	The \undeadref{draugr} is aware of this limitation, and will avoid moisture when not under a \practitionerpossessive{necromancy} direct control.
	
	\undeadrefplural{draugr} are often used to guard ancient tombs, sealed inside where water cannot intrude.
}

\undead{Sea-Draugr}{Sea-Draugar}{sea-draugr}{
	The \undeadrefplural{sea-draugr} are, in many ways, the complete opposite of the regular \undeadrefplural{draugr}.
	Creatures of seas and lakes, they can only be animated from the corpses of those who died by drowning.
	They revel in their water-bloated flesh, lurking beneath the surface and dragging their prey to join them in their watery grave.
	
	A \undeadref{sea-draugr} is subject to the same rules as a regular \undeadref{draugr}.
	It also gains a swimming speed equal to its land speed, or half its flying speed, whichever is greater.
	
	However, instead of remaining dry, an \undeadref{sea-draugr} must remain soaked.
	It functions best when immersed in water, and begins to weaken about five minutes after it emerges.
	After about ten minutes, it dries out too much and completely loses animation.
	
	Regular wetting can extend this time; it might get half an hour in rain, or even an indefinite time if the rain is sufficiently torrential.
	\emph{Continuous} attention using \featref{willing-water-vapour}, or a couple of uses of \featref{willing-water-vapour-2} every 5 minutes, also suffices.
	However, it must be completely immersed for at least 8 hours each day.
	Just like a \undeadref{draugr}, a \undeadref{sea-draugr} is aware of this limitation, and will seek out water.
	
	As long as the corpse remains animated as a \undeadref{sea-draugr}, and sufficiently wetted, it will not continue to rot.
}

\undead{Skeleton}{Skeletons}{skeleton}{
	A \undeadref{skeleton} is the result of reanimating only the bones of a creature, the flesh rotted or carved away.
	The bones arrange themselves in the air, supported by nothing but the will of the animating witch, and the creature's conviction in its own shape.
	The result is a creature far less clumsy than a zombie, but not so resilient.
	
	A corpse reanimated as a \undeadref{skeleton} loses 2 points from all attributes except \attref{grace} and \attref{will}.
	Its \statref{shock-threshold} is also reduced by 2, in addition to the loss from the reduced \attref{might}.
	The mere bones of wings are not sufficient to allow it to fly, if it previously could.
	It also sinks in water, but may move along the bottom.
	
	Requiring only the bones, a \undeadref{skeleton} is not affected by most {\damage} sustained by the corpse.
	Only a critical success on a {\damagetest}, or an intentional effort after death, will typically have broken any bones.
	Likewise, it is not affected by {\damage} in the course of its undeath; any blow insufficient to scatter it across the floor is insufficient to scratch its bones.
	
	A \undeadref{skeleton} lasts a long time without decomposing; at least a decade, and even longer if kept dry.
}

\undead{Living Fossil}{Living Fossils}{fossil}{
	A \undeadref{fossil} is much like a \undeadref{skeleton}, except that the bones have been mineralised, impregnated with stone.
	The essence of earth permeates the creature, strengthening it.
	
	A \undeadref{fossil} uses the same rules as a \undeadref{skeleton}, except that its \attref{might} is not reduced.
	Additionally, fossilised bones do not decay, lasting millennia and more.
}

\undead{Blazing Skeleton}{Blazing Skeletons}{fire-skeleton}{
	A \undeadref{fire-skeleton} can only be made from the charred bones of a creature that died burning.
	Flames race across its bones as it walks, and its eye sockets blaze like the sun.
	It spreads destruction wherever it steps, leaving fire and ash in its wake.
	
	A \undeadref{fire-skeleton} mostly uses the same rules as a \undeadref{skeleton}, with a handful of differences.
	Firstly, the bones used to animate must be charred by fire.
	This should not be enough to totally destroy the bones, but even badly fire-damaged bones have no ill effects on the resulting \undeadref{fire-skeleton}.
	
	Secondly, the \undeadref{fire-skeleton} always burns, as long as it is animated.
	It burns without fuel, and without damaging itself---in fact, it is immune to all harm from fire and heat.
	The fire goes out if the \undeadref{fire-skeleton} loses animation.
	Conversely, the \undeadref{fire-skeleton} loses animation if the fire is extinguished, such as by being immersed in water.
	The fire is somewhat robust, however; it can survive moderate rain, simply causing the droplets to boil away.
	
	The \undeadrefpossessive{fire-skeleton} flames produce heat, and can combust things, just like normal flame.
	They will ignite most combustible materials they touch.
	A \undeadrefpossessive{fire-skeleton} {\unarmed} attacks also {\ignite} the target, at \dice{2}, or add 1 die of {\fire} to a target who is already burning.
}

\undead{Shade}{Shades}{shade}{
	While it is truly an undead, a reanimated corpse, a \undeadrefplural{shade} bears some resemblance to {\ghosts}.
	It shares some of their insubstantial nature, being a part of this realm, but not truly a creature of it.
	It always appears to be wreathed in shadow, and every part of it is dark: black or grey.
	Its facial features are indistinct, or even absent.
	
	Although it is formed from a creature's body, a \undeadref{shade} is insubstantial.
	It often finds its fingers passing straight through objects, like shadows flitting over them.
	Although this makes it harder to affect the world, it also affords the \undeadref{shade} a degree of protection.
	Swords can pass right through it, without even disturbing it.
	
	Light, however, brings the \undeadref{shade} form, clarity.
	This makes it vulnerable.
	Worse still, sunlight can burn it away entirely, destroying it.
	
	A \undeadref{shade} loses 2 points from its \attref{ken}, \attref{charm}, and \attref{presence} scores.
	Furthermore, its insubstantial nature causes it to lose 5 points from its \attref{might} score.
	However, it suffers no penalties to vision in low-light conditions, or even complete darkness.
	And, in complete darkness, it is immune to all physical harm.
	
	Light, even dim light, makes the \undeadref{shade} vulnerable again.
	If the light falls only on part of its body, only that part is vulnerable.
	Sunlight, however, is worse.
	The \undeadref{shade} suffers a \dice{5} {\damagetest} every {\round} that it is exposed to direct sunlight.
	Reducing exposure can reduce the number of dice rolled for the {\damagetest}, but even if it wrapped entirely in thick black cloth, leaving just its eyes exposed so that it might see causes it to suffer a \dice{1} {\damagetest} every round.
	
	Reanimating a corpse as a \undeadref{shade} requires that it is unrotted, but it does not continue to rot while it is animated.
	When the \undeadref{shade} is deanimated, it leaves the corpse fully corporeal again, albeit with a slightly dark pallor, and still affected by any {\damage} the \undeadref{shade} suffered.
	If the \undeadref{shade} is deanimated in sunlight, however, it burns away entirely, leaving no corpse---not even ash.
}

\undead{Wraith}{Wraiths}{wraith}{
	A \undeadref{wraith} is the invention of a foul \practitioner{necromancy} from a bygone era.
	She sacrificed dozens of people to the darkness of a \creatureref{stygian-nightshade}, then recovered their flayed corpses for reanimation.
	The result was a variety of \undeadref{shade} that carried the \creaturerefpossessive{stygian-nightshade} wicked claws, able to rend flesh despite their intangibility.
	The process has been refined since, and needs nothing more than a sprig of \creatureref{stygian-nightshade}.
	However, it still only works on the corpses of creatures that died violent deaths.
	
	A \undeadref{wraith} appears just like a \undeadref{shade}, and uses all the same rules, except for two differences.
	Firstly, the \undeadref{wraith} can affect the physical world with full force; it loses no \attref{might}.
	
	Secondly, it sprouts wicked claws of stygian darkness, increasing the number of dice it rolls for {\unarmed} {\damagetests}.
	A creature without an effective attack gains one, and rolls 2 dice, while a creature with an existing attack rolls at least 3 dice.
	A human, or other creature with proper hands, rolls 5 dice.
}

\undead{Haunt}{Haunts}{souled}{
	%TODO: Souled, Haunts, or something else?
	A \undeadref{souled} is the result of necromancy that is beginning to lift itself from mere reanimation towards the ideal of resurrection.
	It is the result of imbuing a {\soul} into a more conventionally reanimated undead such as a \undeadref{zombie} or \undeadref{skeleton}.
	It is subject to the usual modifications to its statistics, as appropriate to the kind of reanimation.
	
	However, a \undeadref{souled} retains its memories, identity, and free will, and is not subject to the usual hunger.
	It is not controlled by the \practitioner{necromancy} who reanimated it.
	Furthermore, its \attref{might} and \attref{grace} are the only attributes subject to change; the other six are always unchanged.
	It retains all its skills and feats.
	It is still subject to all the benefits and detriments of its loss of biological processes, such as immunity to suffocation, disease, and potions.
	Lastly, it is still subject to usual rules for {\damage} and rotting, so may require \featref{undead-repair}.
	
	A creature can only be reanimated as a single \undeadref{souled} at a time, even when using feats such as \featref{undead-part}.
	
	Names for \undeadrefplural{souled} vary considerably, with many \undeadrefplural{souled} themselves finding the term unpleasant.
	They may refer to themselves as the Souled, or using some other name.
}

\section{Embalming}
\seclabel{embalming}

Decomposition can be such a pain for a \practitioner{necromancy}, putting valuable corpses to waste.
Most corpses barely last more than a week before they are too rotted to make some kinds of undead, such as \undeadrefplural{zombie} and \undeadrefplural{ghoul}.
A \practitioner{necromancy} can always strip away the flesh and raise \undeadrefplural{skeleton}, but these might not suit her needs.
Instead, she might turn to {\embalming}.

Anyone can attempt to {\embalm} a corpse.
The process is a mixture of surgery, and treatment with substances that slow decomposition.
Various substances can be used, with varying effectiveness.
Soaking a corpse in strong alcohol can preserve it for a month or more.
Drying it with salt can preserve it indefinitely, as long as it is not wetted again.
Very long periods of preservation can be achieved with dedicated \featref{embalming-fluid}.

\capital{\embalming} a corpse typically takes a few hours, and requires a \testtypespeciality{ken}{crafting}{Embalmer} Test.
Failure means that the corpse, or some parts of it, won't be preserved, or at least won't last as long as they could.
Particularly bad results can cause {\damage} to corpse.

\capital{\embalming} does nothing to repair {\damage} to the corpse, or to reduce rotting that has already occurred.
It only slows or prevents further rotting.

\section{Souls}
\seclabel{souls}

\capital{\discref{necromancy}} deals with death, after the {\soul} has been separated from the body.
Reanimation uses only the body, while resurrection seeks to reunite body and {\soul}.
Another branch of \discref{necromancy}, however, deals with bare {\souls}---{\ghosts}.

Every living creature has a {\soul}, of some sort.
The more intelligent, reasoning, and self-aware the creature, the stronger the soul.
Truly sapient creatures---humans---are referred to as {\strongsouled}.
Other animals are {\weaksouled}.
The {\souls} of plants, fungi, are the like as so weak as not to exist for any practical purpose---they do not even pass as {\weaksouled}.

Familiars are a particularly unusual case.
They share a shard of their witch's {\soul}, and so are considered to be {\strongsouled} for most purposes.
But some magics affect them strangely, due to their shared soul.

The strength of a {\soul} is a double-edged sword to a \practitioner{necromancy}.
\capital{\strongsouls} are much more substantial and easier to work with for a \practitioner{necromancy} using {\ghosts}.
But they also maintain a link to their own body, brought about by their self awareness.
This makes any body that bore a {\strongsoul} in life much harder to reanimate---the {\soul} must be severed from it first, or it will interfere with the animating energies.

During life, the {\soul} remains in the body.
And at the point of death, it departs the body, and passes through the veil to the afterlife.
Unless something goes wrong.
Sometimes, a {\soul} becomes stuck on this side of the veil, trapped in the mortal realm as a {\ghost}.
\capital{\practitioners{necromancy}} can facilitate this, preventing a {\soul} from passing over, or even reaching through the veil to pull a {\soul} back.

\subsection{Ghosts}
\seclabel{ghosts}

\capital{\ghosts} are {\souls} that have been unable to cross the veil to the afterlife, and remain trapped in the mortal realm.
Some are the intentional creation of \practitioners{necromancy}, magically bound to this realm.
But some also occur naturally.

When a creature, especially a {\strongsouled} creature, dies with unfinished business, it may remain in this realm as a {\ghost}.
The more important the unfinished business---often a matter of vengeance---the more likely a {\ghost} is to form, and the more powerful it is likely to be.
It will remain as long as the business that trapped it remains unfinished.
When the business is finally resolved, it can cross over the veil to the afterlife.
The GM can use these natural {\ghosts} in creating stories, whether or not there is a \practitioner{necromancy} in the coven.

As a creature's {\soul}, a {\ghost} retains most of the capabilities it had in life.
It keeps all attributes and skills.
It is still the same person it was in life---it retains its personality and all its memories, and has total free will.
Even {\ghosts} created by \practitioners{necromancy} are not bound to the \practitionerpossessive{necromancy} will.
However, a {\ghost} that has been drawn back across the veil has no memories of the afterlife, beyond a vague recollection that it is more pleasant than the mortal realm.
It is aware that time has passed, but doesn't know how long.

Despite retaining its skills, personality, and free will, a {\ghost} loses almost all ability to affect the world.
It cannot lift, carry, or move anything; it cannot exert any force on any objects or creatures.
To most people, it is invisible and inaudible, unable to be sensed---it doesn't even appear in the {\mentalrealm}.
People might be able to feel some slight presence, but often won't even realise it represents a {\ghost}.

A few people do learn to sense {\ghosts}, and a few {\ghosts} learn to be sensed by everybody.
To those people who can see ghosts, they appear much as they did in life, except partially transparent, and in muted colours.
However, they tend to appear as they saw themselves, not as how they actually were.
As such, human {\ghosts} tend to appear clothed, wearing their favourite clothes, or the clothes they believed typified them.
It is this need for self-conception that causes {\strongsouls} to form {\ghosts} far more often than {\weaksouls}.

Much as {\ghosts} can't affect objects, they aren't affected by most objects.
They can walk through walls, and pass through other objects and creatures---except other {\ghosts}.
They will not pass through the ground or floor, though, even if they want to, and must still walk upon such a surface.
In water, they can swim much as they did in life.
And, although they can pass through walls, magical barriers like the \featref{circle-contain} will still impede them.

\capital{\ghosts} are also immune to almost everything.
They cannot be harmed by anything except another {\ghost}, or magic explicitly intented to affect {\ghosts}.
They do not need to eat, breathe, drink, or sleep.
They cannot be affected by any potions, poisons, diseases, or the like.
They still suffer {\exhaustion}, as well as {\damage} and {\shock} from the rare things that \emph{can} harm them.
They recover from {\damage} and {\exhaustion} as normal, except that they do not need to sleep, just to rest.
If a {\ghost} is in some way harmed to the point of death, their {\soul} is permanently and irrevocably destroyed.

The inability to affect the world also means that {\ghosts} of witches lose most ability to work magic.
They lose the benefit of all feats, for all intents and purposes.
They have not lost the knowledge required, but the world simply refuses to respond to their will.

Note that the above rules describe a typical {\ghost}: the weakest kind, the kind a novice \practitioner{necromancy} might create.
Some {\ghosts} become more powerful, through help from a \practitioner{necromancy}, from sheer effort of will, or from learning to affect the world once again.
The GM may add other abilities to naturally-occurring {\ghosts} as they see fit, possibly using the {\ghost} related feats in this chapter as inspiration.

\subsection{Phylacteries}
\seclabel{phylacteries}

Resurrection, even to the limited extent that it is possible, requires the return of the creature's {\soul}.
While an experienced \practitioner{necromancy} might reach through the veil between worlds to pluck the {\soul} from whatever afterlife it may be enjoying, it can be much easier to keep the {\soul} shackled to the mortal realm.
Such is the purpose of a {\phylactery}.

A {\phylactery} is an object into which a shard of a person's {\soul} has been bound, enchanted so that the rest of the {\soul} will join it when it would otherwise pass on.
\capital{\phylacteries} must be created from a clay jar, at least the size of a fist but possibly larger.
They are no more robust than the jars they are created from, and their destruction frees the shard of {\soul} within.
The destruction of a {\phylactery} is always felt by the person whose {\soul} it contains, just as the death of a familiar, but otherwise carries no ill effects.
If the {\phylactery} is destroyed after the person has died and their entire {\soul} has passed into it, their {\soul} is released to pass on to the afterlife.

A person can have no more than one {\phylactery} at a time, and the previous one must always be destroyed before a new one can be created.
Likewise, a single object cannot be the {\phylactery} for more than one person at a time.

It should be noted that a {\phylactery} does nothing to \emph{prevent} a person's death; it only makes it easier to restore them afterwards.
However, as long as a witch's {\soul} remains in the {\phylactery} and does not depart this realm, the witch's death does not kill her familiar.

\section{Feats}

\feat{Raise Zombie}{animate-zombie}{20}{
	\noprereq
}{
	You can restore a terrible facsimile of life to the bodies of deceased animals, reanimating it as a \undeadref{zombie}.
	For now, you are limited to animals at least as large as a mouse, and no larger than a medium-size dog such as a bloodhound.
	You can't manage any creature that was {\strongsouled} in life, due to interference from the link that remains.
	
	\materials{An animal corpse, a \circleref{small}, a lit candle which the ritual extinguishes}
	
	The reanimation ritual takes five minutes, and must be performed in the dark.
}

\feat{Raise Skeleton}{animate-skeleton}{15}{
	\featref{animate-zombie}
}{
	After a few reanimations, most \undeadrefplural{zombie} are starting to come apart at the seams a bit.
	There comes a time when it's easier just to strip all the flesh off and make the bones stand up by themselves.
	You may reanimate the bones of an animal corpse as a \undeadref{skeleton}, subject to the same limitations as \featref{animate-zombie}.
	
	\materials{The bones of an animal corpse (with the flesh removed), a \circleref{small}, a lit candle which the ritual extinguishes}
	
	The reanimation ritual takes five minutes, and must be performed in the dark.
}

\feat{Raise Ghoul}{animate-ghoul}{20}{
	\skillref[1]{necromancy},
	\featref{animate-zombie}
}{
	\undeadrefplural{ghoul} are faster and scarier than \undeadrefplural{zombie}, but also \emph{hungrier}.
	You may reanimate an animal corpse as a \undeadref{ghoul}, subject to the same limitations as \featref{animate-zombie}.
	
	\materials{An animal corpse, an additional corpse to be consumed by the \undeadref{ghoul}, a \circleref{small}, a lit candle which the ritual extinguishes}
	
	The reanimation ritual takes five minutes, and must be performed in the dark.
	At the conclusion of the ritual, the newly-arisen \undeadref{ghoul} must immediately be fed a complete corpse---of the same kind of animal as the ghoul---or it does not fall under the \practitionerpossessive{necromancy} control.
}

\feat{Raise Draugr}{animate-draugr}{20}{
	\skillref[1]{necromancy},
	\featref{animate-zombie}
}{
	By animating corpses as \undeadrefplural{draugr}, you can keep them around longer than mere \undeadrefplural{zombie}.
	You may reanimate a desiccated animal corpse as a \undeadref{draugr}, subject to the same limitations as \featref{animate-zombie}.
	
	\materials{An animal corpse {\embalmed} by desiccation (drying out), a \circleref{small}, a lit candle which the ritual extinguishes}
	
	The reanimation ritual takes five minutes, and must be performed in the dark.
}

\feat{Raise Sea-Draugr}{animate-sea-draugr}{10}{
	\skillref[1]{necromancy},
	\featref{animate-draugr}
}{
	A small variation on the ritual to animate \undeadrefplural{draugr} lets you create the opposite.
	You may reanimate a drowned, soaked animal corpse as a \undeadref{sea-draugr}, subject to the same limitations as \featref{animate-zombie}.
	
	\materials{A water-soaked animal corpse that died by drowning, a \circleref{small}, a lit candle which the ritual extinguishes}
	
	The reanimation ritual takes five minutes, and must be performed in the dark.
	The corpse must be kept soaked during the ritual, so keep a few buckets of water handy.
	And make sure the \materialref{ritual-circle} won't be washed away.
}

\feat{Raise Fossil}{animate-fossil}{20}{
	\skillref[1]{necromancy},
	\featref{animate-skeleton}
}{
	Fossilisation is naturally a slow process, but a dedicated \practitioner{necromancy} can accelerate the process.
	You may reanimate the bones of an animal corpse as a \undeadref{fossil}, subject to the same limitations as \featref{animate-zombie}.
	
	\materials{The bones of an animal corpse (with or without flesh) buried in a bog, a \circleref{small}, a small heap of finely crushed rock, a lit candle which the ritual extinguishes}
	
	Beginning the reanimation ritual requires five minutes, but the \undeadref{fossil} does not rise for 24 hours.
	For the entire 24 hours, the candle must remain lit, the \materialref{ritual-circle} must remain intact, and the area must remain dark.
	The witch need not be present for the whole duration, however.
	
	Over the course of the 24 hour period, the rock dust is drawn into the bog and incorporated into the bones, and any remaining flesh rots away.
	At the conclusion, the \undeadref{fossil} is animated and claws its way to the surface.
	The witch must be present at this point to assert immediate control over the \undeadref{fossil}, or it becomes uncontrolled.
}

\feat{Raise Blazing Skeleton}{animate-fire-skeleton}{20}{
	\skillref[2]{necromancy},
	\featref{animate-skeleton}
}{
	Playing with fire is dangerous, and playing with \undeadrefplural{fire-skeleton} is even worse.
	But you've decided it's worth the risk.
	You may reanimate the charred bones of an animal that died burning as a \undeadref{fire-skeleton}, subject to the same limitations as \featref{animate-zombie}.
	
	\materials{The charred bones of an animal that died burning (with the flesh removed), a \circleref{small}, a lit candle which the ritual extinguishes}
	
	The reanimation ritual takes five minutes, and must be performed in the dark.
	The ritual ignites the bones, but it requires fire to do it.
	Normally the candle suffices, but if the candle is substituted for a {\phylactery}, a flame must still be provided.
}

\feat{Raise Shade}{animate-shade}{20}{
	\skillref[1]{necromancy},
	\featref{animate-zombie}
}{
	A \undeadref{shade} is a valuable tool in a \practitionerpossessive{necromancy} arsenal, silent and deadly in the dark.
	You may reanimate an animal corpse as a \undeadref{shade}, subject to the same limitations as \featref{animate-zombie}.
	
	\materials{An animal corpse, a \circleref{small}, a lit candle which the ritual extinguishes}
	
	The reanimation ritual takes five minutes, and must be performed in the dark, \emph{at night}.
}

\feat{Raise Wraith}{animate-wraith}{20}{
	\skillref[2]{necromancy},
	\featref{animate-shade}
}{
	Using the victim of a violent death, and a sprig of \creatureref{stygian-nightshade} placed in its mouth, you can raise a powerful and violent variety of \undeadref{shade}: a \undeadref{wraith}.
	You may reanimate an animal corpse as a \undeadref{wraith}, subject to the same limitations as \featref{animate-zombie}.
	
	\materials{The corpse of an animal which died violently, \herbcreature{stygian-nightshade}{5}, a \circleref{small}, a lit candle which the ritual extinguishes}
	
	The reanimation ritual takes five minutes, and must be performed in the dark, \emph{at night}.
}

\feat{Deanimate}{deanimate}{10}{
	\featref{animate-zombie}
}{
	You can withdraw the animating force from a creature you have reanimated, returning it to death.
	This is far neater than beating it to death.
	
	\materials{An undead under your control, a \materialref{ritual-circle} of the same size required to initially animate the undead, an unlit candle which the ritual lights}
	
	The ritual takes five minutes, and must be performed in a brightly lit location.
	The undead must remain within the \materialref{ritual-circle} for the duration.
	You can affect multiple undead with one casting of the ritual, as long as they are in within the \materialref{ritual-circle} for the duration.
	The size of the \materialref{ritual-circle} required is determined by the size of the largest undead in the group.
}

\feat{Uncontrolled Deanimation}{deanimate-2}{10}{
	\featref{deanimate}
}{
	You can use the \featref{deanimate} ritual against even an uncontrolled undead.
}

\feat{Offensive Deanimation}{deanimate-3}{10}{
	\skillref[1]{necromancy},
	\featref{deanimate}
}{
	You can use the \featref{deanimate} ritual against any undead, be it under your control, under another \practitionerpossessive{necromancy} control, uncontrolled, or even a \undeadref{souled}.
}

\feat{Maintain Control}{undead-control}{10}{
	\featref{animate-zombie}
}{
	Keeping control of your undead can pose a real challenge, especially when you need to sleep.
	You've learned a ritual to place them in a sort of suspended animation, keeping them ready to reassert control over when you awake.
	
	\materials{An undead under your control, a \materialref{ritual-circle} of the same size required to initially animate the undead, a lit candle}
	
	The ritual takes five minutes, and the undead must remain within the \materialref{ritual-circle} for the duration of casting the ritual.
	When you have finished casting the ritual, the undead enters suspended animation; it simply remains in the \materialref{ritual-circle}, unmoving.
	It still counts against the number of undead you are controlling, but you need not remain nearby, or conscious.
	At any point when you are within range, you may end the ritual's effect and reassert control over the undead.
	
	The effect of the ritual lasts as long as the candle remains lit.
	With a large enough candle, you can get 2 or 3 days out it.
	If the candle is extinguished, or removed, you must be in a position to immediately resume active control of the undead---otherwise it becomes uncontrolled.
	
	If you have the ability to control multiple undead, you may affect multiple undead with one \materialref{ritual-circle} and one invocation of the ritual.
	The size of the \materialref{ritual-circle} required is determined by the size of the largest undead in the group.
	You may end the effect of the ritual on each undead individually.
}

\feat{Assert Control}{undead-control-2}{10}{
	\skillref[1]{necromancy},
	\featref{undead-control},
	\featref{deanimate-2}
}{
	Rather than drawing the animation out of undead, you can simply substitute your own animating force, taking control of the undead.
	This uses the \featref{deanimate} ritual, but requires an additional material: a lit candle which the ritual extinguishes---the same as animating an undead in the first place.
	
	On the completion of the ritual, instead of losing animation, the undead comes under your control.
	Undead you take control of this way still count against the maximum number of undead you can control, and exceeding this limit will free an earlier undead from your control, just as raising a new one would.
	
	Obviously, this is useless against an undead you already control, but it is useful in combination with \featref{deanimate-2} or \featref{deanimate-3}.
	You can even take control of a type of undead that you could not normally raise in this way.
	However, you cannot use this against undead that cannot normally be subject to a \practitionerpossessive{necromancy} control, such as \undeadrefplural{souled}---you must simply deanimate these.
}

\feat{Precision Control}{control-deanimate-small}{10}{
	\skillref[1]{necromancy},
	\featref{undead-large},
	\featref{undead-control} or \featref{deanimate}
}{
	You can use a \circleref{small} for the \featref{undead-control} or \featref{deanimate} rituals---assuming you have the feat to perform the ritual at all---regardless of the size of circle you would require to animate the creature in the first place.
	Note, however, that the undead must fit inside the circle, so you will need slightly bigger than a \circleref{small} for an elephant, or the like.
}

\feat{Rapid Control}{control-deanimate-fast}{20}{
	\skillref[2]{necromancy},
	\featref{undead-control} or \featref{deanimate}
}{
	You have become far faster at manipulating an undead's animating force---it can hardly be called a ritual anymore, though it still requires a \materialref{ritual-circle}.
	You can perform the \featref{undead-control} or \featref{deanimate} rituals in just one {\action}.
}

\feat{Contact Control}{control-deanimate-fast-2}{20}{
	\skillref[3]{necromancy},
	\featref{control-deanimate-small},
	\featref{control-deanimate-fast}
}{
	You can perform the \featref{undead-control} or \featref{deanimate} rituals without a \materialref{ritual-circle}, simply by touching the undead you want to affect.
	You still need all the requisite candles, however.
	Furthermore, you can only affect one undead per action if you do not use a \materialref{ritual-circle}.
}

\feat{Stitches}{undead-repair}{10}{
	\featref{animate-zombie}
}{
	Many reanimations and resurrections are ineffective on corpses which are too badly damaged.
	By sealing wounds, stitching severed parts back on, and gluing bones together, you can solve this.
	Any parts you reattach must come from the original creature.
	
	The repair and reanimation requires a Test, with the {\tn} determined by how badly damaged the corpse is, using your choice of \skillref{necromancy} or \skillref{healing}.
	A successful Test repairs at least enough {\damage} to restore the creature's \statref{shock-threshold} to 1, and a high result may repair even more.
	You must perform the repairs while the corpse is dead; you cannot repair it while it is animated.
}

\feat{Scraps}{undead-repair-2}{10}{
	\skillref[1]{necromancy},
	\skillref[1]{healing},
	\featref{undead-repair}
}{
	You can do more than stitch a damaged corpse back together; you can stitch \emph{several} corpses together.
	When using \featref{undead-repair}, you may assemble the corpse to be animated out of parts from different corpses.
	A corpse assembled out of several individually intact parts can be healthier than a single, damaged corpse.
	
	The pieces must all come from creatures of the same kind; all from humans, all from dogs, and so on.
	They must be assembled to form a creature of that kind; you cannot make a six-legged dog.
}

\feat{Chimera}{undead-repair-3}{25}{
	\skillref[3]{necromancy},
	\skillref[2]{healing},
	\featref{undead-repair-2}
}{
	You have mastered the art of assembling corpses, creating horrifying, chimeric monstrosities.
	When using \featref{undead-repair-2}, the parts needn't all come from the same kind of animal.
	They needn't form a normal creature, either; you could stitch extra legs on to a dog.
	
	The creature can typically use replacement anatomy easily; for example, if you replace a human's arms with a bear's.
	New anatomy, however---an extra pair of limbs, for example---may take several hours, or even days to learn.
	Neither a human with an animal mouth, nor an animal with a human mouth, can speak properly.
	
	The GM may invent a set of statistics for the resulting creature, based upon the component creatures and modified, as usual, by its kind of undeath.
}

\feat{Darning}{undead-repair-active}{15}{
	\skillref[1]{necromancy},
	\skillref[2]{healing},
	\featref{undead-repair}
}{
	You may make repairs to a corpse even while it is currently animated.
	Any Tests made to do so use \skillref{healing}.
	You may even reattach severed parts, though these must be the original parts unless you also have \featref{undead-repair-2}.
	\capital{\featref{undead-repair-3}} even allows you to attach parts from different kinds of creature.
}

\feat{Knitted Resurrection}{undead-repair-active-2}{15}{
	\skillref[2]{necromancy},
	\skillref[3]{healing},
	\featref{undead-repair-active}
}{
	You have discovered a means to resurrect dead tissue by attaching it to living tissue.
	This allows you to reattach severed parts to a person or animal.
	Any Tests made to do so use \skillref{healing}.
	
	This doesn't do much, if anything, to heal {\damage}; no more than normal surgery.
	The reattached parts, however, become living parts of the creature, for all intents and purposes.
	Some rot---up to about a week---is tolerable, though disgusting, and will be healed naturally after reattachment.
	
	The reattached parts must be the original parts, unless you also have \featref{undead-repair-2}.
	If you do have \featref{undead-repair-2}, however, you may replace failed organs, or broken limbs, with healthy ones from another creature.
	The target must remain alive throughout the entire process, so replacing a heart is incredibly difficult, and replacing a brain is impossible.
	\capital{\featref{undead-repair-3}} even allows you to attach parts from different kinds of creature.
}

\feat{Major Undead}{undead-large}{20}{
	\skillref[1]{necromancy},
	\featref{animate-zombie}
}{
	Larger bodies need more force to reanimate, but it's force you've learned to provide.
	When you perform a {\reanimationritual}, you may use a \circleref{medium} instead of a \circleref{small}, in order to ignore the upper size limit on the creature.
	You still cannot reanimate a {\strongsouled} creature, such as a human; you need \featref{undead-human} to do so.
}

\feat{Undead Head}{undead-head}{10}{
	\featref{animate-zombie}
}{
	Rather than bothering to reanimate larger creatures, you can just reanimate parts of them.
	The head, specifically, the seat of consciousness.
	
	You may reanimate a creature's severed head using a {\reanimationritual}.
	The usual restrictions apply; for example, you cannot reanimate a human unless you also have \featref{undead-human}.
	However, when evaluating whether you need \featref{undead-large} and a \circleref{medium}, consider only the size of the creature's head, not the whole creature.
	As such, any head short of an elephant's only needs a \circleref{small}.
	
	The resulting creature has all the limitations you would expect from a severed head.
	It can't move, and can only bite people who put their hands in its mouth.
	It has no \attref{might} or \attref{grace} scores for most purposes, though it retains its \attref{might} score for calculating its \statref{shock-threshold}, and for biting.
	It can still see, hear, and so on, and vocalise or speak as it could in life.
}

\feat{Partial Undead}{undead-part}{15}{
	\skillref[1]{necromancy},
	\featref{undead-head}
}{
	Sometimes, a \practitioner{necromancy} has to make do with just the parts of corpses that they can find.
	Sometimes, of course, they just want a hand.
	
	You may reanimate just part of a creature using a {\reanimationritual}.
	As with \featref{undead-head}, you cannot reanimate a human without \featref{undead-human}, and whether you need to use \featref{undead-large} is determined by the size of the part you reanimate.
	However small the part you reanimate, the result still counts as one undead under your control.
	
	The abilities of a reanimated part are determined by which part it is, and the GM should adjudicate this according to common sense.
	For example, a part without the head has no sense of sight, hearing, smell, or taste; only touch.
	Similarly, a single limb is likely to move more slowly than a full creature, and will suffer a considerable penalty to attacks---if it can attack at all.
	\capital{\statref{dodge-rating}} is also likely to be reduced.
	\capital{\statref{shock-threshold}} and \statref{resilience} are unlikely to be affected, however.
	Furthermore, the {\damage} the resulting undead part has sustained is determined only by the {\damage} to the relevant part, and it is not considered to have taken any {\damage} merely by being severed from the rest of the body.
	
	If you reanimate a part without a mouth, this causes problems when the creature breaks free of your control and becomes hungry.
	It cannot eat, so is behaviour will become increasingly erratic.
	It may begin to simply hunt and stockpile food, or something else, at the GM's discretion.
	Partial \undeadrefplural{ghoul} are even worse; if they cannot be fed, they cannot be controlled at all.
}

\feat{Sever Soul}{undead-human}{20}{
	\skillref[1]{necromancy},
	\featref{undead-large} or \featref{undead-head}
}{
	Reanimating a creature that was once {\strongsouled} has previously been impossible, due to interference from the residual link.
	You've learned to sever these links, and hence reanimate these creatures.
	
	You may reanimate a {\strongsouled} creature, such as a human, or an animal that was a familiar, using a {\reanimationritual}.
	You must use an iron blade as part of the ritual, to sever the link.
	
	Reanimating an entire human requires a \circleref{medium}, and \featref{undead-large}.
	Reanimating smaller parts of a human, using \featref{undead-head} or \featref{undead-part}, may not.
	
	A reanimated familiar has lost the link to its witch, and is now just a normal animal of its kind.
	See \featref{reanimate-familiar} to reanimate your familiar without losing this link.
}

\feat{Undead Familiar}{reanimate-familiar}{10}{
	\featref{animate-zombie}
}{
	While a {\soul} normally interferes with reanimating a creature, you've begun to figure out how to use it to your advantage, beginning on the path towards resurrection.
	Unfortunately, you can't actually summon any {\souls} back to their bodies yet.
	Not to worry, though, for you have quite ready access to one {\soul} in particular: your familiar's, so inextricably bound to your own.
	
	If your familiar dies and you can recover the corpse, you can reanimate it, paying no XP cost beside that required to purchase this feat in the first place.
	You may use any {\reanimationritual}, and it becomes the appropriate kind of \undeadref{souled}.
	
	Reanimating a familiar in this way does not prevent recovering it through the usual repetition of the binding ritual later (see the section \secref{familiar-injury-death}), although the normal XP cost must still be paid each time that method used.
}

\feat{Pull Yourself Together!}{skeleton-reassemble}{20}{
	\skillref[2]{necromancy},
	\skillref[1]{healing},
	\featref{animate-skeleton},
	\featref{undead-repair-active},
	\featref{undead-part}
}{
	By animating each bone of a \undeadref{skeleton} separately, you give it the ability to reassemble itself when scattered.
	When you animate a \undeadref{skeleton}, \undeadref{fossil}, or \undeadref{fire-skeleton}, you may give it this ability.
	This can make it quite hard to deanimate, however, so bear in mind that you don't have to.
	
	If an undead with this ability is destroyed by a {\damagetest}, or some other trauma that scatters its bones without destroying them, it does not lose animation.
	Over the next few minutes, the bones will draw themselves back together and reform the undead creature, which will begin normal operation again.
	The bones can be restrained to prevent this, but even so, will keep trying to draw themselves together, indefinitely.
	Even a \practitioner{necromancy} in control of the undead cannot give an order to prevent this; it is unresponsive to orders until reassembled.
	
	Thankfully, this does not entirely prevent deanimating the skeleton.
	It will lose animation if enough of the bones are broken, or, in the case of a \undeadref{fire-skeleton}, extinguished.
	Alternatively, the \featref{deanimate} ritual still works.
}

\feat{Undead Phoenix}{skeleton-reassemble-phoenix}{20}{
	\skillref[3]{necromancy},
	\skillref[1]{healing},
	\featref{skeleton-reassemble},
	\featref{animate-fire-skeleton}
}{
	You have learned an obscure ritual to emulate the legendary phoenix, creating an avian undead that is reborn from fire in an instant.
	When you use \featref{animate-fire-skeleton} on the bones of a flying bird, you may and use this feat.
	If you do so, the resulting undead gains all the effects of \featref{skeleton-reassemble}, as well as several extras.
	
	Firstly, despite its skeletal nature, the bird can fly just as well as it could in life.
	Secondly, its flame burns unstoppably; it cannot be extinguished in any way, except by deanimating it.
	Lastly, it can reassemble itself faster, taking only moments, not minutes.
	After it is destroyed, it reforms on its next {\turn}, though it cannot do anything else on that {\turn}.
	
	However, it is said that the phoenix is a unique bird---that only one ever exists at a time.
	While not quite true here, \emph{you} can certainly only create one.
	The last one you created must lose animation before you can animate another one---it is not enough to simply lose control of it.
}

\feat{Phylactery}{phylactery}{10}{
	\skillref[1]{necromancy},
	\featref{reanimate-familiar},
	\featref{undead-large} or \featref{undead-head}
}{
	You've learned to restore your familiar's {\soul} to its body in the event of its death.
	The next step is simply to perform the same procedure upon \emph{yourself}.
	This is complicated by the fact that you are dead, of course, so you ought to have a very good plan in place for pulling this off.
	Examples include a resurrection pact with a trustworthy friend who knows this same procedure, having your familiar do it (\featref{phylactery-familiar}), or ensuring you can stick around to do it yourself (\featref{projection-lifeline-phylactery}).
	
	Firstly, this feat allows you to extract a sliver of your own {\soul} and place it in a {\phylactery}.
	The ritual to do so requires an hour, and must be performed in a dark place.
	It costs 10 XP each time you perform the ritual, as you extract another sliver of your {\soul}.
	
	\materials{The clay jar to become the {\phylactery}, a drop of your own blood, \herb[deadly nightshade]{belladonna}{2}, a \circleref{medium}}
	
	Secondly, you have learned to use a {\phylactery} in a reanimation ritual.
	This must be the {\phylactery} containing the {\soul} of the person whose body is being reanimated, but you may do this with anybody's {\phylactery}, not just your own.
	
	You may use any {\reanimationritual}, reanimating the person as the appropriate kind of \undeadref{souled}.
	The {\phylactery} takes the place of the candle in the ritual.
	The ritual typically requires a \circleref{medium} and \featref{undead-large}---unless you take advantage of \featref{undead-head}, or \featref{undead-part}.
	You do not need \featref{undead-human}---you are incorporating the {\soul}, not severing it.
}

\feat{Familiar Resurrection}{phylactery-familiar}{10}{
	\skillref[1]{necromancy},
	\featref{phylactery}
}{
	Reanimating yourself is hard, what with being dead and all.
	So you've taught your familiar to do it for you.
	
	Your familiar can perform the {\reanimationritual} variant granted by \featref{phylactery}.
	It is only the link between the {\soul} in your {\phylactery} and the sliver of the same {\soul} in your familiar that affords it the magical intuition to do this, so it can only perform the ritual in order to reanimate \emph{you}.
}

\feat{Self-Sacrifice}{phylactery-lethal}{10}{
	\skillref[1]{necromancy},
	\featref{phylactery}
}{
	Dividing your {\soul} is always costly, making the typical method of creating a {\phylactery} rather draining.
	Fortunately, it is possible to create a {\phylactery} without dividing your {\soul}.
	It's simple, really---you just shift your \emph{entire} {\soul} into the {\phylactery} at once.
	This is, of course, lethal.
	
	\materials{The clay jar to become the {\phylactery}, a deadly dose of \herb[deadly nightshade]{belladonna}{2}, a \circleref{medium}}
	
	The ritual requires only a minute, and must be performed in a dark place.
	It has no XP cost.
	It kills your body and traps your entire {\soul} in the {\phylactery}.
	
	If you have \featref{projection-start-2} or a \featref{projection-potion}, and an alternative {\lifeline}, you can leave your body in time that your mind survives the ritual.
	If you have \featref{projection-lifeline-phylactery}, you may even transition seamlessly to using the {\lifeline} provided by the {\phylactery} you are creating.
}

\feat{Touching the Veil}{death-detection}{10}{
	\noprereq
}{
	When a {\soul} departs our world for the next, its passage disrupts the veil between worlds.
	A witch who knows what to look for can feel this disruption.
	
	You can feel where people have died, though this sense is damped by both distance and time.
	If you pass through the actual position of the death, you'll notice for up to about two weeks after it occurred.
	You automatically sense a death in the same room only for a few days after it's happened, and in the same house for only about a day.
	However, a Test can reveal slightly older or more distant deaths, if you are searching for them.
	You can't gain any information about the identity of the victim or the cause of death.
	
	Locations of mass or repeated death can leave their traces lingering for much longer.
	The site of a battlefield or sacrificial altar may be felt for many years after.
}

\feat{Ghost Sight}{ghost-detect}{10}{
	\noprereq
}{
	To most people, {\ghosts} are invisible, inaudible, almost entirely entirely undetectable.
	But not to you.
	You can always see, hear, and even smell {\ghosts}.
	Still, however, you cannot touch them, and they cannot touch you.
	
	You can block out this sense, if a {\ghost} is annoying you.
	However, you cannot block out a {\ghost} that would be detectable even without this feat.
}
