\storychapter{Mental Preparation}

Bonnie ran her finger over the book, straining to read her mentor's scratchy handwriting.
She muttered the words to herself.
``{\dots}and bring to boil. Soak jasmine flower to count of ten, then remove.''
She cast a glance at the cauldron, where bubbles were just beginning to run up the sides of the vessel.
The flower lay dried on the bench already---she'd prepared the ingredients earlier---and she grabbed it in the large tongs.
Into the cauldron, and start the count.

``One potato, two potato, three potato, four pota---''

\storybreakword{SLAM}

Something smashed into her mind.
Her vision blurred.
The tongs tumbled forward into the boiling water as her hands went to her head.

Quickly, \emph{think}.
Mistress Evenny had taught her how to resist this.
Imagine a fortress, walls of unyielding stone.
Build it around your mind, but don't hide within.
Stand prou---

\storybreakword{THUNK}

Everything was{\dots}\ Not grey, because grey meant sight, and there was no sight here.
But it was dull.
Filled with fog, but not the fog of a winter morning.
It was the fog of a clouded mind, and it stretched away in all directions.
There was no ground, no sky.
Just the fog.

The Mindplace.
That was what Mistress Evenny called it.
Bonnie hated it here.

She tried to recall her training.
\emph{This} was her.
That was the first thing to remember.
There was nothing here that she could sense, but that didn't matter.
\emph{She} was here.
Always was, and always would be.
And she ought never to forgot that, or she'd have lost herself.

Next{\dots}\ That thing---that light---beside her.
That was her body, left behind.
It'd be dim; there was nobody in there now, after all.
Except{\dots}\ It wasn't.
Why not?

Oh!
There was someone \emph{else} in there, of course.
She nearly slapped herself on the forehead, except she didn't have hands or a face right now.
Someone, or some\emph{thing}, had possessed her.
It'd be running around in her body right now, doing goodness knows what with it.

Well, she couldn't be having with that!
She grabbed ahold that mind that now resided where her own should be, and \emph{pushed}, trying to force her way back in.
It doesn't yield an inch.
Until, with a sudden release of pressure, it did.

\storybreakword{POP}

She was back in her own body.
\emph{Stocktake}.
She was back by the bench with the herbs.
Whoever'd been in her body had made a mess of it.
Everything she'd been using had been swept aside, but someone had started another pile of herbs in its place.
It looked{\dots}\ Nefarious.
Hemlock, bloodwort, wolfsbane.
Something else she didn't recognise.
Someone must be trying to poison her!

She went to sweep them off the bench, but stopped when she saw the writing on the wall.
Just behind the bench, beside the shelves of herbs, someone had scorched words onto the wood of the cottage's wall.

``FETCH SILVER''

It was signed, in the same smouldering charcoal: ``EVENNY''.

Oh!
Oh.
Bonnie sighed.
Right.
Of course.
Dutifully, she turned and ran for her bedroom.

She grabbed the jewellery box from beneath the bed, and dumped it out on the floor, then snatched her necklace from the pitiful pile.
It was a shell, yellow and red and a beautiful spiral shape, from all the way out near the sea.
But the chain was silver.
Mostly.
She was going to be \emph{so annoyed} if Mistress Evenny broke it.

She felt a pressure against her mind as she ran back to the kitchen.
It wasn't anything close to the terrible blows of last time, thankfull.
Mistress Evenny was being polite, this time.
She sighed, and let go.

\storybreakword{SHOVE}

The Mindplace again.
She was patient now, waiting until Mistress Evenny was done with{\dots} Whatever she was doing.
It only seemed a few seconds, in the end, before the bright thing---Mistress Evenny's mind, she knew---suddenly shot off and streaked away into the distance.
She plunged herself back into the dull glow left behind.

\storybreakword{PLOP}

Argh!
\emph{Bitter}.
She nearly spat the wad of herbs out of her mouth, but she caught sight of more writing burned into the wall.
``CHEW''.
Screwing up her face against the taste, she obeyed.

Her mind raced as her teeth pulped the plants, squeezing their juices into her mouth.
Wolfsbane?
Hemlock?
\emph{Bloodwart?}
She really hoped Mistress Evenny knew what she was doing, or Bonnie wasn't going to live to see the sunrise{\dots}

She'd only been chewing about ten seconds when that mental pressure returned, and she ceded control of her body again{\dots}

\storybreakword{SHUNT}

Another minute in the Mindplace and{\dots}

\storybreakword{WHOA}

Bonnie stumbled backwards, away from the edge.
She was on the \emph{roof?}
And a bottle in her hand, which she'd ben holding out precariously over the edge.
No instructions---Mistress Evenny probably hadn't wanted to burn something into the thatch.
So she probably ought to get back to how she'd been left{\dots}

She shuffled back towards the edge, and stretched her arm out into the air, dangling the bottle.
Just in time, too, as a black speck appeared in the sky over the horizon, growing quickly larger as it skimmed over the forest canopy.
Bonnie stretched, holding the bottle as far away from herself as possible{\dots}

\emph{Whoosh}.
A gust of air knocked her flat on her back---and nearly off the roof---as the bottle was snatched from her hand.
Mistress Evenny barely lost any speed as she banked sharply in the air above the cottage and shot back the way she'd come.
Bonnie heaved a sigh of relief.
She'd have to ask what \emph{that} had all been about later.

But for now, she had to get this bitter taste out of her mouth.
Come to think of it, she probably ought to take an emetic.
\emph{Ergh}.
First, though, she had to get down{\dots}
