Bonnie ran her finger over the book, straining to read her mentor's scratchy handwriting.
She muttered the words to herself.
``{\dots}and bring to boil. Soak hemlock flower to count of ten, then remove.''
She cast a glance at the cauldron, where bubbles were just beginning to run up the sides of the vessel.
The hemlock flower lay dried on the bench already---she'd prepared the ingredients earlier---and she grabbed it in the large tongs.
Into the cauldron, and start the count.

``One potato, two potato, three potato, four pota---''

\storybreakword{SLAM}

Something smashed into her mind.
Her vision blurred.
The tongs tumbled forward into the boiling water as her hands went to her head.

Quickly, \emph{think}.
Mistress Evenny had taught her how to resist this.
Imagine a fortress, walls of unyielding stone.
Build it around your mind, but don't hide within.
Stand prou---

\storybreakword{THUNK}

Everything was{\dots}\ Not grey, because grey meant sight, and there was no sight here.
But it was dull.
Filled with fog, but not the fog of a winter morning.
It was the fog of a clouded mind, and it stretched away in all directions.
There was no ground, no sky.
Just the fog.

The Mindplace.
That was what Mistress Evenny called it.
Bonnie hated it here.

She tried to recall her training.
\emph{This} was her.
That was the first thing to remember.
There was nothing here that she could sense, but that didn't matter.
\emph{She} was here.
Always was, and always would be.
And she ought never to forgot that, or she'd have lost herself.

Next{\dots}\ That thing---that light---beside her.
That was her body, left behind.
It'd be dim; there was nobody in there now, after all.
Except{\dots}\ It wasn't.
Why not?

Oh!
There was someone \emph{else} in there, of course.
She nearly slapped herself on the forehead, except she didn't have hands or a face right now.
Someone, or some\emph{thing}, had possessed her.
It'd be running around in her body right now, doing goodness knows what with it.

Well, she couldn't be having with that!
She grabbed ahold that mind that now resided where her own should be, and \emph{pushed}, trying to force her way back in.
It doesn't yield an inch.
Until, with a sudden release of pressure, it did.


